### 2022-10-26 draft_report.Rmd
10/26 updates

---
title: "Untitled"
output: html_document
date: '2022-10-18'
---

### 1. Introduction
two different methods of haplotype reconstruction (haploqa vs qtl2)
  haploqa was written a decade ago, not well-maintained
  qtl2 is newer method. Since haploqa is not maintained, it's ideal to discontinue. However, before the transition is made, this project is in place
  to compare the results to ensure the qtl2 outputs can be used as a substitute for the haplotype reconstruction tasks that haploqa is doing.

```{r}
source(paste0(root,"/input_data_prep_functions.R"))
```

### 2. Methods

#### 2.1. Environment Set-up

The calculation is being done for 5 samples in GigaMUGA (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), and the MiniMUGA sample. A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all the relevant file paths associated with each cross. A preview of the configuration file is shown as below:

```{r}
root <- dirname(rstudioapi::getSourceEditorContext()$path)
config <- fread(paste0(root, '/annotations_config.csv'))
config
```

Below is an example of the pipeline for CC:
```{r}
sample_type <- 'CC'
### Environment
config_sample <- config[config$array_type == sample_type]
marker_type <- config_sample$marker_type
founders_list <- unlist(strsplit(config_sample$founders_list, ", "))
n_founders <- length(founders_list)
ngen <- config_sample$ngen
sample_url <- config_sample$url
founder_url <- config_sample$founder_url

# data output directory
data_dir <- file.path(root, config_sample$data_dir)
```

#### 2.2. Data Retrieval

The pipeline takes the according data_dir, or qtl2_dir values for each cross, concatenate the variable values with the working directory (which is the directory where the functions are stored in), and read in/output files from the directories.

Input data was retrieved from haploqa.jax.org, and converted into the required input format for qtl2. There are two parts of data wrangling tasks:
1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports automatically for each individual within each cross
2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input file for qtl2

For example, for cross 'Collaborative Cross', the URL to the sample is http://haploqa-dev.jax.org/tag/Collaborative%20Cross.html

The 'sample_summary_scrape' function reads the html elements of the website, (?)

```{r}
## summary file
haploqa_html <- read_html(sample_url)
# extract information from html file
html_temp <- haploqa_html %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_temp[grepl('/sample', html_temp)])
summary_df <- sample_summary_scrape(haploqa_html, url_list, marker_type)

## individual samples
### this downloads the data directy to the folder, 'data_dir'. Try not to actually run this loop
haploqa_html <- read_html(sample_url)
# extract information from html file
html_temp <- haploqa_html %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_temp[grepl('/sample', html_temp)])

# list of urls to generate samples from
url_ind <- unique(summary_df$`Sample Filepath`)

# individual samples
inc = 0
for (url in url_ind) {
  inc = inc + 1 # increment
  file <- sample_individual_scrape(url, url_domain)
  print(paste0('Working on file ', inc, '/', length(url_ind), ': ', file))
  sample_df_save <- as.data.frame(content(GET(file)))
  print('Writing to directory')
  file_name <- unlist(strsplit(file, '/'))[6]
  GET(file, write_disk(paste0(data_dir, '/', file_name), overwrite = TRUE), show_col_types = FALSE)
}

### note: get_qtl2_input reads all files from data_dir recursively and binds them into one big file. The big file is then merged with summary file to gather some necessary metadatas. This merged file is the main input file

```

The main input file, which contains (insert detailed info here later), is then processed into the required input format for qtl2.
The qtl2 input file contains the following parts: genotype, genetic map, ... (do we need to explain thw input file structures to the audience or would they be familiar with this already?)

Encoding(?) - in the main file, the retrieved data can be separated into two parts:
    1. variables 'haplotype1', 'haplotype2', which represents the haploqa results,
    2. variables 'allele1', 'allele2', which is the (original?) alleles used to calculate the haploqa results. These alleles will also be used for the qtl2 computations.

Function 'get_qtl2_input' contains all the necessary algorithms to convert the data into separate required files. The output is then read in by qtl2 and used for calculations for different metrics.

  1. qtl2 result calculations (genoprob(pr), phased geno(ph_geno), etc)
  2. haploqa result calculations, with the same metrics as above.

#### 2.3. Demo (?)

##### 2.3.1. GigaMUGA
The pipeline for GigaMUGA executes all the individuals within the sample. The contributing founders for all individuals within the GigaMUGA samples are always the same.

```{r}
# simulated phenotypes
list_pheno <- c('WBC', 'NEUT')
```

Results for DO:
```{r}
# DO
do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)

```

Below are the genotype visualizations for some individuals: 
(pick some representative ones?)
```{r}
plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

Results for CC:
```{r}
# DO
cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)

```

genotype visualizations for some individuals: 
```{r}
plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

(do this for other models as well)

##### 2.3.2. MiniMUGA
MiniMUGA has a separate pipeline that runs for one individual at a time due to each individuals having different (contributing?) founders.

```{r}
sample1_results <- sample_haplotype_reconstruction('MiniMUGA', 'DJN', samples_gen = T, qtl2_file_gen = T)
plot_onegeno_test(sample1_results[['ph_geno']], sample1_results[['ph_geno_haploqa']], sample1_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
sample2_results <- sample_haplotype_reconstruction('MiniMUGA', 'JXN', samples_gen = T, qtl2_file_gen = T)
plot_onegeno_test(sample2_results[['ph_geno']], sample2_results[['ph_geno_haploqa']], sample2_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

##### 2.3.3. Shiny Visualizations

```{r}
source(paste0(root,"/app.R"))
shinyApp(ui, server)
```

#### 3.Statistical Results

There are three parts of the comparison:
1. haploqa results, which can be obtained directly from haploqa.jax.org by downloading the SNP level haplotype reports from each sample. 
2. qtl2 genAIL results, which was calculated using the R package 'qtl2' with the crosstypes 'genail'
3. qtl2 best model, which was calculated using the qtl2package with the original crosstype

The visualization shows a high-level idea of the difference between the haplotype reconstruction results between qtl2 and haploqa. To compare the detailed results for the methods, ...

(where to put truth model)

#### 4. Conclusion - TBD

### 2022-10-26 HaploQA_qtl2_output.R

@@ -20,79 +20,44 @@ library(stringi)
 library(lsa)
 library(readr)
 
-### BXD: cross_info - all 1s, choose one founder genome to be A. Make sure foundergeno and cross_info have consistent labels for genomes.
-# geno for founders - http://haploqa-dev.jax.org/tag/UNC_Villena_GIGMUGV01_20141012_FinalReport.html
-# pick a color for DBA
-#
-# 1. run BXD
-# 2. dynamic cross_info instead of hard-coding
-# 3. number of crossovers, visualize in some way
-# 4. make cos_sim plots on high level (mean of a sample across entire genome)
-# 5. have Shiny work for BXD as well
-# 6. modularize and write documentations for functions
-# 7. http://haploqa-dev.jax.org/tag/f2.html
-
-
 # source the function script
 root <- dirname(getSourceEditorContext()$path)
 source(paste0(root,"/input_data_prep_functions.R"))
 
+# simulated geno
 list_pheno <- c('WBC', 'NEUT')
 
+### GigaMUGA pipelines
 do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-
-do_truth_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = T)
-
-## count pct of missing values in sample geno
-
 cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-cc_truth_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = T, samples_gen = F, truth_model = T)
-
 bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-bxd_truth_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F, truth_model = T)
-
[…]

### 2022-10-26 app.R

@@ -9,20 +9,20 @@ source(paste0(root,"/input_data_prep_functions.R"))
 results_dir <- file.path(root, 'results')
 rds_dir <- file.path(results_dir, 'RDS')
 
-total_ind <- 277 # different for CC and DO - fix it
-
+# phased geno files
 phased_geno_comp_cc <- shiny_viz_input('CC', rds_dir) # contains phased geno of qtl2 AND haploqa
 # contains: qtl2, haplo, map
 phased_geno_comp_do <- shiny_viz_input('DO', rds_dir)
-
 phased_geno_comp_bxd <- shiny_viz_input('BXD', rds_dir)
-
 phased_geno_comp_f2 <- shiny_viz_input('F2', rds_dir)
 
+# comparison files
+truth_comp_fp <- file.path(results_dir, 'shiny_pct_csvs')
 
+# minimuga data
 minimuga_dir <- '/Users/linc/Documents/GitHub/HaploQA_qtl2_comp/minimuga_qtl2_genail/'
 
-
+### add a legend that shows which side is which method
 ### to launch shiny app
 ui <- fluidPage(
   titlePanel("Geno-wide Genotype comparison - qtl2 vs. HaploQA"),
@@ -32,25 +32,34 @@ ui <- fluidPage(
     sidebarPanel(width = 3,
                  #htmlOutput("selectUI"),
                  #selectInput("individual", "Individual", ''),
-                conditionalPanel(
+                 conditionalPanel(
                    condition = "input.select == 'Diversity Outbred'",
                    selectInput("individual", "Individual", choices = seq(1:277), selected = 1)),
                  conditionalPanel(
                    condition = "input.select == 'Collaborative Cross'",
-                   selectInput("individual1", "Individual", choices = seq(1:222), selected = 1)),
+                   selectInput("individual1", "Individual", choices = seq(1:218), selected = 1)),
                  conditionalPanel(
                    condition = "input.select == 'BXD F3'",
                    selectInput("individual2", "Individual", choices = seq(1:71), selected = 1)),
-                conditionalPanel(
-                  condition = "input.select == 'F2'",
-                  selectInput("individual3", "Individual", choices = seq(1:34), selected = 1)),
-                conditionalPanel(
-                  condition = "input.select == 'MiniMUGA'",
-                  selectInput("individual4", "Individual", choices = seq(1:25), selected = 1))
+                 conditionalPanel(
+                   condition = "input.select == 'F2'",
[…]

### 2022-10-26 input_data_prep_functions.R

@@ -384,6 +384,9 @@ get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, s
   ### combine all files
   df_raw <- rbindlist(lapply(data_files, read_sample_txt)) %>% filter(!sample_id %in% exclude_list) # save Y chrom here for gender plotting
   
+  ## exclude the ones with bad string names
+  summary_df <- summary_df %>% filter(grepl('CC', `Strain Name`)) 
+  
   # clean up summary df
   sum_df <- summary_df %>% 
     select(ID, Sex, `% Het. Calls`, `% No Call`, Platform) %>% rename(sample_id = ID)
@@ -392,7 +395,9 @@ get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, s
   
   # eliminate those that has no call > 10%
   ### use checkifnot to make sure the markers/SNP are in the annotation file (annotation file is the boss)
-  df_all <- merge(df_raw, sum_df, by = 'sample_id') %>% filter(`% No Call` < 10) %>% filter(!chromosome %in% c('0', 'Y', 'M')) %>% filter(!sample_id %in% exclude_list) 
+  df_all <- merge(df_raw, sum_df, by = 'sample_id') %>% 
+    filter(`% No Call` < 10) %>% filter(!chromosome %in% c('0', 'Y', 'M')) %>% 
+    filter(!sample_id %in% exclude_list) 
   stopifnot("There are markers in qtl2 dataframe not present in annotation file, check validity or exclude such markers" = any(!unique(df_all$snp_id) %in% annot_df$marker) == FALSE) # if there is any marker from df that's not in annotation, raise error
   
   
@@ -899,7 +904,8 @@ get_geno_pct_diff <- function(ginf_haploqa, ginf_qtl2, summary_df, num_chr, foun
 #
 # @return results (list of dataframes) - a list containing all objects necessary for computations
 haplotype_reconstruction_pipeline <- function(sample_type, list_pheno, qtl2_file_gen, samples_gen, truth_model) {
-  #sample_type <- 'DO'
+  #sample_type <- 'CC'
+  #truth_model = T
   ## results directory
   results_dir <- file.path(root, 'results')
   dir.create(results_dir, showWarnings = FALSE) 
[…]

### 2022-10-26 qtl2_temp.R
Delete qtl2_temp.R

@@ -1,136 +1,0 @@
-library(shiny)
-library(qtl2)
-library(dplyr)
-library(data.table)
-library(rstudioapi)
-library(stringi)
-library(ggplot2)
-
-### requirements:
-# 1. config file
-# 2. gmap 
-# 3. summary_df
-
-# source the function script
-root <- dirname(getSourceEditorContext()$path)
-source(paste0(root,"/input_data_prep_functions.R"))
-
-# config file
-config <- fread(paste0(root, '/annotations_config.csv'))
-## results directory
-results_dir <- file.path(root, 'results')
-dir.create(results_dir, showWarnings = FALSE) 
-# to save rds
-rds_dir <- file.path(results_dir, 'RDS')
-dir.create(rds_dir, showWarnings = FALSE) 
-
-all_results_gen = T
-num_chr <- c((1:19),"X")
-
-#### RDS functions
-sample_types <- c('CC', 'DO') # right now supports CC, DO, GigaMUGA, MiniMUGA
-results_all_file <- file.path(rds_dir, 'results_all.rds')
-if(all_results_gen == T) {
-  list_results_all <- rds_loop(sample_types, config, results_dir, rds_dir, num_chr) 
-  print('saving final results to RDS file')
-  saveRDS(list_results_all, file = results_all_file)
-} else {
-  list_results_all <- readRDS(results_all_file)
-}
-
-cc_results <- list_results_all[['CC']]
-do_results <- list_results_all[['DO']]
-
-## list components: summary, data_dir, cross, map (same for qtl2 and haploqa)
-## pr, ginf, ph_geno, pos (qtl2 only)
-## ginf_haploqa, ph_geno_haploqa
-
-## attributes from cc_test
-total_ind <- 277 # take this automatically from cc_test?
-
-##############
-
-# lookup tables
-founder_codes <- c('AA', 'BB', 'CC', 'DD', 'EE', 'FF', 'GG', 'HH')
-single_codes <- sapply(founder_codes, function(x) substr(x, 1, 1))
-map_codes <- c(1, 2, 3, 4, 5, 6, 7, 8)
-founder_lookup <- setNames(founder_codes, map_codes)
-founder_reverse <- setNames(map_codes, single_codes)
-founders_dict <- fread(paste0(root, '/founder_strains_table.csv'))
[…]

### 2022-10-27 HaploQA_qtl2_output.R
updated functions 10/27

@@ -24,20 +24,27 @@ library(readr)
 root <- dirname(getSourceEditorContext()$path)
 source(paste0(root,"/input_data_prep_functions.R"))
 
+# results directory
+results_dir <- file.path(root, 'results')
+dir.create(results_dir, showWarnings = FALSE) 
+
 # simulated geno
 list_pheno <- c('WBC', 'NEUT')
 
 ### GigaMUGA pipelines
-do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
+do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
+cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
+bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
 
 ## GigaMUGA truth models
-do_truth_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = T)
-cc_truth_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = T, samples_gen = F, truth_model = T)
-bxd_truth_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F, truth_model = T)
-f2_truth_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = F, truth_model = T)
+do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = T, samples_gen = F)
[…]

### 2022-10-27 input_data_prep_functions.R

@@ -609,32 +609,35 @@ save_rds <- function(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_fo
   founder_lookup_table <- fread(file.path(root, 'founder_lookup_table.csv'))
   founder_all_rev_lookup <- setNames(founder_lookup_table$founder_codes, founder_lookup_table$founder_id)
   
-  # simulate maxmarg to calculate haploqs ginf 
-  ginf_haploqa_fp <- paste0(rds_dir, "/ginf_haploqa_", sample_type, ".rds")
-  if (!file.exists(ginf_haploqa_fp)) {
-    print('haploqa ginf file does not exist, running calculations from maxmarg simulation')
-    ginf_haploqa <- maxmarg_sim(summary_df, data_dir, num_chr, founder_all_rev_lookup, cross = cross_list[['cross']], qtl2_dir, sample_type, n_founders, founder_haplo_lookup)
-    ## save to rds file
-    cross_list[['ginf_haploqa']] <- ginf_haploqa
-    saveRDS(ginf_haploqa, file = ginf_haploqa_fp)
-  } else {
-    print('haploqa ginf file exists, reading in file')
-    ginf_haploqa <- readRDS(ginf_haploqa_fp)
-    cross_list[['ginf_haploqa']] <- ginf_haploqa
+  if (truth_model == F) { # only need HaploQA for genail models
+    # simulate maxmarg to calculate haploqs ginf 
+    ginf_haploqa_fp <- paste0(rds_dir, "/ginf_haploqa_", sample_type, ".rds")
+    if (!file.exists(ginf_haploqa_fp)) {
+      print('haploqa ginf file does not exist, running calculations from maxmarg simulation')
+      ginf_haploqa <- maxmarg_sim(summary_df, data_dir, num_chr, founder_all_rev_lookup, cross = cross_list[['cross']], qtl2_dir, sample_type, n_founders, founder_haplo_lookup)
+      ## save to rds file
+      cross_list[['ginf_haploqa']] <- ginf_haploqa
+      saveRDS(ginf_haploqa, file = ginf_haploqa_fp)
+    } else {
+      print('haploqa ginf file exists, reading in file')
+      ginf_haploqa <- readRDS(ginf_haploqa_fp)
+      cross_list[['ginf_haploqa']] <- ginf_haploqa
+    }
+    
+    # simulate guess_phase for haploqa to calculate phased geno
[…]

### 2022-10-27 draft_report.Rmd
Update draft_report.Rmd

@@
 
+HaploQA is a web application developed in 2015 by the Jackson Laboratory for the purpose of storing, increasing accessibility and quality assurance of micro array mice genotyping data.
+According to the HaploQA about website, 'On the import of a sample, the haplotype probabilities that are represented in the plots are calculated.' and the haplotype reconstruction process for the contributing founder strains is then done based on the calculated probabilities.
+
+R/qtl2 is an open-source software package built in 2019, which performs similar haplotype reconstruction tasks as mentioned in HaploQA - the software takes genotyping data as input, calculate the haplotype probabilities, then perform haplotype reconstructions based on the quantitative results.
+(HMM)
+
+Since R/qtl2 was built more recently and is currently maintained by the developers, the Jackson Laboratory is considering transitioning from HaploQA to R/qtl2 for similar haplotype reconstructing tasks with genotyping data.
+This project contains pipelines that retrieve genotyping data used in the initial HaploQA algorithm, performs haplotype reconstructions with said data using R/qtl2, and compare the R/qtl2 results to the initial results generated in HaploQA to determine whether R/qtl2 would be an appropriate replacement method for the HaploQA web application.
+
+(Explain metrics used to assess model - percentage of markers that agree with the optimal model)
+
+### 2. Methods
+
+The pipeline developed for this project is stored in the script, 'input_data_prep_functions.R', which is stored in the same working directory as this report and is sourced below:
+
 ```{r}
@@
 
-### 2. Methods
+in order to decide whether to transition, haploqa and qtl2 each need to be compared to an optimal model.
+
+, this project will perform haplotype reconstruction tasks using genortping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), 70 (?) individuals within the MiniMUGA sample.
+Decision wa made to not do Muga/MegaMUGA due to the discontinued usage.
 
@@
 
-The calculation is being done for 5 samples in GigaMUGA (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), and the MiniMUGA sample.
 A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all the relevant file paths associated with each cross.
 A preview of the configuration file is shown as below:
+
+(explain the config file, contains crosstype, marker annotations, path to directories, etc.)
 
@@
 config <- fread(paste0(root, '/annotations_config.csv'))
-config
+print(config)
 ```
 
-Below is an example of the pipeline for CC:
+Below is an example of how the pipeline set-up looks like for CC:
+
 ```{r}
+# comment it!
+#
 sample_type <- 'CC'
@@
 
-Input data was retrieved from haploqa.jax.org, and converted into the required input format for qtl2. There are two parts of data wrangling tasks:
-1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports automatically for each individual within each cross
-2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input file for qtl2
+Input data was retrieved from haploqa.jax.org, and converted into the qtl2 input format. The data wrangling consists of two parts:
+1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports for each individual within each cross
+2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input files for qtl2
 
@@
 # individual samples
+## 
 inc = 0
@@
 
-The visualization shows a high-level idea of the difference between the haplotype reconstruction results between qtl2 and haploqa. To compare the detailed results for the methods, ...
+##### 3.1. Truth models
 
-(where to put truth model)
+At a glance of the visualizations which present a high-level idea of the difference between the haplotype reconstruction results between qtl2 and haploqa, there are some visible differences between the two models - specifically, qtl2 appears to have more crossovers and come minor areas appear to have different founder results.
+To determine which model is more correct, some 'truth' models, which the algorithms have been used (for a long time?), were also constructed for each of the crosses.
+
+As mentioned in the configuration file, the truth model for CC uses crosstype 'risib8', the truth model for DO uses crosstype 'do', the truth model for BXD F3 uses crosstype 'ail', and the truth model for F2 uses crosstype 'f2'.
+The executions of pipelines for the truth models are as shown below:
+
+```{r}
+do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
+cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
+bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
+f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
+```
+
+##### 3.2. Quantitative comparisons
+Combining the results of the truth model, the qtl2 haplot
+(Display the csvs, and perhaps the visualizations here too)
+
+```{r}
+# filepath to the csvs
+shiny_pct_fp <- file.path(results_dir, 'shiny_pct_csvs')
+
+# DO
+do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
+do_pct_df
+```
+
+```{r}
+# CC
+cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
+cc_pct_df
+
+```
+
+Based on the existing results for qtl2 and haploqa, and truth model result comparisons, it can be observed that...
 
 #### 4. Conclusion - TBD
+
+#### 5. References
+1. http://haploqa-dev.jax.org/help.html
+
+2.
+Broman KW, Gatti DM, Simecek P, Furlotte NA, Prins P, Sen Ś, Yandell BS, Churchill GA.
+R/qtl2: Software for Mapping Quantitative Trait Loci with High-Dimensional Data and Multiparent Populations.
+Genetics.
+2019 Feb;211(2):495-502.
+doi: 10.1534/genetics.118.301595.
+Epub 2018 Dec 27.
+PMID: 30591514; PMCID: PMC6366910.

### 2022-10-27 draft_report.Rmd
10/27 fixed minor issues

@@
 ### 1. Introduction
-two different methods of haplotype reconstruction (haploqa vs qtl2)
-  haploqa was written a decade ago, not well-maintained
-  qtl2 is newer method. Since haploqa is not maintained, it's ideal to discontinue. However, before the transition is made, this project is in place
-  to compare the results to ensure the qtl2 outputs can be used as a substitute for the haplotype reconstruction tasks that haploqa is doing.
 
@@
 This project contains pipelines that retrieve genotyping data used in the initial HaploQA algorithm, performs haplotype reconstructions with said data using R/qtl2, and compare the R/qtl2 results to the initial results generated in HaploQA to determine whether R/qtl2 would be an appropriate replacement method for the HaploQA web application.
+
+(some backgrounds about when the optimal models were built/what about them makes them optimal?) Some optimal models were built for each of the crosses in qtl2, which
 
@@
 cc_pct_df
+#cc_pct_df
+ggplot(cc_pct_df) + aes(x = qtl2_pct_diff, y = haploqa_pct_diff) + geom_point()
+```
 
+```{r}
+# BXD f3
+bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
+bxd_pct_df
+
+```
+
+```{r}
+# BXD f3
+f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
+f2_pct_df
+
+```
+
+```{r}
+do_qtl2_truth_diff <- do_pct_df$qtl2_pct_diff
+cc_qtl2_truth_diff <- cc_pct_df$qtl2_pct_diff
+bxd_qtl2_truth_diff <- bxd_pct_df$qtl2_pct_diff
+f2_qtl2_truth_diff <- f2_pct_df$qtl2_pct_diff
+
+# align lengths
+max_n <- max(nrow(do_pct_df), nrow(cc_pct_df), nrow(bxd_pct_df), nrow(f2_pct_df))
+
+length(do_qtl2_truth_diff) <- max_n 
+length(cc_qtl2_truth_diff) <- max_n
+length(bxd_qtl2_truth_diff) <- max_n
+length(f2_qtl2_truth_diff) <- max_n
+
+df_qtl2_truth_diff <- as.data.frame(cbind('do_qtl2_truth_diff' = do_qtl2_truth_diff, 'cc_qtl2_truth_diff' = cc_qtl2_truth_diff, 'bxd_qtl2_truth_diff' = bxd_qtl2_truth_diff, 'f2_qtl2_truth_diff' = f2_qtl2_truth_diff))
+df_qtl2_truth_diff
+```
+
+```{r}
+do_qtl2_truth_diff
 ```

### 2022-10-27 app.R

@@ -1,5 +1,6 @@
 library(rstudioapi)
 library(shiny)
+library(data.table)
 
 
 root <- dirname(getSourceEditorContext()$path)

### 2022-10-27 input_data_prep_functions.R

@@ -1713,13 +1713,13 @@ geno_align <- function(df, n_founders) {
 geno_all_comp <- function(sample_type, sample_results, truth_results) {
   num_chr <- c((1:19),"X")
   #sample <- '6UY'
-  #sample_type <- 'CC'
-  #sample_results <- cc_results
-  #truth_results <- cc_truth_results
-  #founder_lookup_table <- fread(file.path(root, 'founder_lookup_table.csv'))
-  #founder_all_rev_lookup <- setNames(founder_lookup_table$founder_codes, founder_lookup_table$founder_id)
+  #sample_type <- 'BXD'
+ # sample_results <- bxd_results
+  #truth_results <- bxd_truth_results
+  founder_lookup_table <- fread(file.path(root, 'founder_lookup_table.csv'))
+  founder_all_rev_lookup <- setNames(founder_lookup_table$founder_codes, founder_lookup_table$founder_id)
   
-  geno_codes <- colnames(cc_truth_results[['pr']]$X)
+  geno_codes <- colnames(truth_results[['pr']]$X)
   founder_all_rev_lookup <- setNames(geno_codes, seq(1, length(geno_codes)))
   
   config <- fread(paste0(root, '/annotations_config.csv'))
@@ -1764,7 +1764,7 @@ geno_all_comp <- function(sample_type, sample_results, truth_results) {
   df_geno_all_chr <- list()
   
   for (chr in num_chr) {
-    #chr <- 'X'
+    #chr <- 1
     
     print(chr)
     ## raw geno
@@ -2221,6 +2221,7 @@ truth_model_reconstruction <- function(sample_type, list_pheno, qtl2_file_gen, s
   
   ## summary file
   summary_df_fp <- paste0(data_dir, '/', sample_type, '_summary.csv')
+  summary_df <- fread(summary_df_fp)
   
   ## list of samples to exclude, if any
   exclude_list <- unlist(strsplit(config_sample$exclude_samples, ", "))

### 2022-11-28 draft_report.Rmd
updated documentations

@@
 
-HaploQA is a web application developed in 2015 by the Jackson Laboratory for the purpose of storing, increasing accessibility and quality assurance of micro array mice genotyping data.
+HaploQA is a web application developed by Kieth Sheppard for The Reinholdt's Lab in 2015 for the purpose of storing, increasing accessibility and quality controlling of micro array mice genotyping data.
 According to the HaploQA about website, 'On the import of a sample, the haplotype probabilities that are represented in the plots are calculated.' and the haplotype reconstruction process for the contributing founder strains is then done based on the calculated probabilities.
+The source codes of HaploQA was written in Python 3.
 
-R/qtl2 is an open-source software package built in 2019, which performs similar haplotype reconstruction tasks as mentioned in HaploQA - the software takes genotyping data as input, calculate the haplotype probabilities, then perform haplotype reconstructions based on the quantitative results.
-(HMM)
+R/qtl2 is an open-source software package built in 2019, which not only performs similar haplotype reconstruction tasks as HaploQA but also has the functionality to perform other computations such as genetic mapping - in the pipeline developed for this project,the software takes genotyping data as input, calculate the haplotype probabilities based on a Hidden Markov Model algorithm, then perform haplotype reconstructions based on the quantitative results.
 
-Since R/qtl2 was built more recently and is currently maintained by the developers, the Jackson Laboratory is considering transitioning from HaploQA to R/qtl2 for similar haplotype reconstructing tasks with genotyping data.
-This project contains pipelines that retrieve genotyping data used in the initial HaploQA algorithm, performs haplotype reconstructions with said data using R/qtl2, and compare the R/qtl2 results to the initial results generated in HaploQA to determine whether R/qtl2 would be an appropriate replacement method for the HaploQA web application.
+Since R/qtl2 was built more recently and is currently maintained by the developers, the Jackson Laboratory is considering transitioning from HaploQA to R/qtl2 generalized AIL (Advanced intercross lines)  for similar haplotype reconstructing tasks with genotyping data.
+The main purpose of this project is to decide if the transition should be made determine based on whether HaploQA or R/qtl2 genail provides more accurate haplotype reconstruction results.
 
-(some backgrounds about when the optimal models were built/what about them makes them optimal?) Some optimal models were built for each of the crosses in qtl2, which
+The best way to determine which method provides more accurate results is to compare the results of the two methods each to the 'correct' results - in R/qtl2, there exists some cross-specific models which the Hidden Markov model was tailored to generate the most accurate results for a particular cross.
+These models will be referred to as the 'truth' models in this document.
 
-(Explain metrics used to assess model - percentage of markers that agree with the optimal model)
+In this project, we selected some GigaMUGA crosses, developed pipelines that retrieve genotyping data used in the initial HaploQA algorithm, performs haplotype reconstruction computations for both qtl2 genail and haploqa, and compare the percentages of markers within each method that agree with the optimal model of the according crosses to determine whether HaploQA or R/qtl2 gen-AIL would be the most appropriate for similar haplotype reconstruction tasks.
+
+The analysis was also done for MiniMUGA, as it is the standard cross type JAX uses for mutant mice, however, there is no tailored truth model for MiniMUGA.
 
@@
 ```{r}
+
+library(rstudioapi)
+library(qtl2)
+library(dplyr)
+library(tidyverse)
+library(httr)
+library(rvest)
+library(data.table)
+library(qtl2convert)
+library(ggrepel)
+library(reshape2)
+library(stats)
+library(ggplot2)
+library(stringr)
+library(stringi)
+library(lsa)
+library(readr)
+
+root <- dirname(getSourceEditorContext()$path)
 source(paste0(root,"/input_data_prep_functions.R"))
+
+#source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
+
 ```
 
-in order to decide whether to transition, haploqa and qtl2 each need to be compared to an optimal model.
-
-, this project will perform haplotype reconstruction tasks using genortping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), 70 (?) individuals within the MiniMUGA sample.
-Decision wa made to not do Muga/MegaMUGA due to the discontinued usage.
+As mentioned in the introduction, in order to decide whether to transition, the accuracy of haploqa and qtl2 each need to be compared to an optimal model.
+This project will perform haplotype reconstruction tasks using genortping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), 70 (?) individuals within the MiniMUGA sample.
+Decision was made to not include Muga/MegaMUGA data due to the discontinued usage of the array types.
 
@@
 A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all the relevant file paths associated with each cross.
+The config file contains each individual crosstypes, their according marker annotation files, paths to the directory where the input data is stored and the directory where qtl2 input data is saved in, etc.
+
+The purpose of this configuration file is to avoid having to pass a large list of parameters into each pipeline function.
+
 A preview of the configuration file is shown as below:
 
-(explain the config file, contains crosstype, marker annotations, path to directories, etc.)
-
 ```{r}
-root <- dirname(rstudioapi::getSourceEditorContext()$path)
+#root <- dirname(rstudioapi::getSourceEditorContext()$path)
 config <- fread(paste0(root, '/annotations_config.csv'))
@@
 
-```{r} 
+```{r, eval = FALSE} 
 ## summary file
@@
 The qtl2 input file contains the following parts: genotype, genetic map, ... (do we need to explain thw input file structures to the audience or would they be familiar with this already?)
+
+There are three major pipelines:
+a. 'haplotype_reconstruction_pipeline' - haplotype reconstructions for an entire cross (all individuals within the cross) 
+b. 'sample_haplotype_reconstruction' - haplotype reconstructions for one individual 
+c. 'truth_model_reconstruction' - haplotype reconstructions for optimal/truth models
+
+whereas the cross, MiniMUGA, contains mutate mice where all individuals may have different contributing founder strains. Therefore, the MiniMUGA samples
 
@@
 ```
-
-(do this for other models as well)
 
@@
 #cc_pct_df
-ggplot(cc_pct_df) + aes(x = qtl2_pct_diff, y = haploqa_pct_diff) + geom_point()
+#ggplot(cc_pct_df) + aes(x = qtl2_pct_diff, y = haploqa_pct_diff) + geom_point()
 ```
@@
 
-df_qtl2_truth_diff <- as.data.frame(cbind('do_qtl2_truth_diff' = do_qtl2_truth_diff, 'cc_qtl2_truth_diff' = cc_qtl2_truth_diff, 'bxd_qtl2_truth_diff' = bxd_qtl2_truth_diff, 'f2_qtl2_truth_diff' = f2_qtl2_truth_diff))
-df_qtl2_truth_diff
+df_qtl2_truth_diff <- as.data.frame(cbind('do' = do_qtl2_truth_diff, 'cc' = cc_qtl2_truth_diff, 'f3' = bxd_qtl2_truth_diff, 'f2' = f2_qtl2_truth_diff))
+plot_df <- melt(df_qtl2_truth_diff)
+
+library(ggplot2)    
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('qtl2 - proportion concordance between optimal and implemented models')
+```
+
+According to the visualization, which shows the percentage of marker differences between qtl2 genail and the truth models - DO, F3 and F2 are all producing results that are almost exactly the same as the optimal model, with CC having some discrepancies.
+We believe the reason why CC performs so differently is that the optimal model for CC does not allow any heterozygosity, whereas the genail does, so the genail model showed some residual heterozygosity from the CC mice which the optimal model does not have the ability to show.
+
+```{r}
+do_haploqa_truth_diff <- do_pct_df$haploqa_pct_diff
+cc_haploqa_truth_diff <- cc_pct_df$haploqa_pct_diff
+bxd_haploqa_truth_diff <- bxd_pct_df$haploqa_pct_diff
+f2_haploqa_truth_diff <- f2_pct_df$haploqa_pct_diff
+
+# align lengths
+max_n <- max(nrow(do_pct_df), nrow(cc_pct_df), nrow(bxd_pct_df), nrow(f2_pct_df))
+
+length(do_haploqa_truth_diff) <- max_n 
+length(cc_haploqa_truth_diff) <- max_n
+length(bxd_haploqa_truth_diff) <- max_n
+length(f2_haploqa_truth_diff) <- max_n
+
+df_haploqa_truth_diff <- as.data.frame(cbind('do' = do_haploqa_truth_diff, 'cc' = cc_haploqa_truth_diff, 'f3' = bxd_haploqa_truth_diff, 'f2' = f2_haploqa_truth_diff))
+plot_df <- melt(df_haploqa_truth_diff)
+
+library(ggplot2)    
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa - proportion concordance between optimal and implemented models')
+```
+
+According to the visualization, which shows the percentage of marker differences between haploqa and the truth models - the haploqa results for f2 and f3 still seem to be similar to the optimal model, but DO and CC seem to be relatively incorrect results.
+
+```{r}
+do_haploqa_truth_diff <- do_pct_df$haploqa_qtl2_pct_diff
+cc_haploqa_truth_diff <- cc_pct_df$haploqa_qtl2_pct_diff
+bxd_haploqa_truth_diff <- bxd_pct_df$haploqa_qtl2_pct_diff
+f2_haploqa_truth_diff <- f2_pct_df$haploqa_qtl2_pct_diff
+
+# align lengths
+max_n <- max(nrow(do_pct_df), nrow(cc_pct_df), nrow(bxd_pct_df), nrow(f2_pct_df))
+
+length(do_haploqa_truth_diff) <- max_n 
+length(cc_haploqa_truth_diff) <- max_n
+length(bxd_haploqa_truth_diff) <- max_n
+length(f2_haploqa_truth_diff) <- max_n
+
+df_haploqa_truth_diff <- as.data.frame(cbind('do' = do_haploqa_truth_diff, 'cc' = cc_haploqa_truth_diff, 'f3' = bxd_haploqa_truth_diff, 'f2' = f2_haploqa_truth_diff))
+plot_df <- melt(df_haploqa_truth_diff)
+
+library(ggplot2)    
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('proportion concordance between implemented models (haploqa vs qtl2)')
 ```
@@
 ```{r}
-do_qtl2_truth_diff
+mean_df <- rbind(t((colMeans(do_pct_df[,2:4]))), t((colMeans(cc_pct_df[,2:4]))), t((colMeans(bxd_pct_df[,2:4]))), t((colMeans(f2_pct_df[,2:4])))) %>% as.data.frame()
+colnames(mean_df) <- paste(colnames(mean_df), "mean", sep = "_")
+
+mean_df$sample_type <- c('DO', 'CC', 'F3', 'F2')
+mean_df %>% select(sample_type, everything())
+colMeans(cc_pct_df[,2:4])
 ```
 
-Based on the existing results for qtl2 and haploqa, and truth model result comparisons, it can be observed that...
+MiniMUGA
+
+```{r}
+mini_plot <- mini_pct_res %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
+plot_df <- melt(mini_plot)
+plot_df$variable <- 'MiniMUGA'
+
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA - percentage of marker difference between genail and haploqa') + xlab('model')
+
+```
+
+The MiniMUGA samples are mutant mice samples, meaning they often have founders that contribute to only a few markers.
+The above boxplot shows a relatively significant high number of markers that are different between haploQA and qtl2 genail, which shows a weakness for qtl2: due to the way HMM models determines genome states based on probabilities - when they are extra founders, or founders that only contribute to a few markers, qtl2 genail tends to be 'confused' and bounce states in the haplotype reconstruction process, whereas haploQA is capable of staying in states.
 
 #### 4. Conclusion - TBD
+Overall, the existing results for qtl2 genail and haploqa show that the two methods have very similar performance for most of the crosses, with genail performing significantly better for the DO outbred mice samples.
+HaploQA handles the extra founders being present in the haplotype reconstruction process better, however, qtl2 genail model has stronger capabilities to produce more accurate results for complicated crosses.
+So based on the results we have so far, we’re leaning toward that genail is more accurate than haploqa.
 

### 2022-11-28 HaploQA_qtl2_output.R

@@ -3,6 +3,23 @@
 ### annotation config saved in annotations_config.csv
 ### founder strain dictionary saved in founder_strains_table.csv
 
+### output files for haplotype reconstruction
+
+
+## everything that starts from BJSJL
+
+# take one founder out (cast), use genail 7
+# 6UY DO
+# run reconstruction, plot onegeno and run error lod
+# find in python where haploqa generates the plots
+
+# 1. DO genail 7, do error lod computations
+# 2. document
+# 3. haploqa codes
+# 4. 
+sample_res <- sample_haplotype_reconstruction('MURGIGV01', 'PX', samples_gen = F, qtl2_file_gen = T)
+
+
 library(rstudioapi)
 library(qtl2)
 library(dplyr)
@@ -19,6 +36,10 @@ library(stringr)
 library(stringi)
 library(lsa)
 library(readr)
+
+# 2, 4, 8, 16 founders
+# pipeline for one GigaMUGA mutant sample
+# system.time()
 
 # source the function script
 root <- dirname(getSourceEditorContext()$path)
@@ -44,36 +65,226 @@ bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen
 f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = T, samples_gen = F)
 
 
-qtl2::plot_onegeno(cc_truth_results[['ph_geno']], cc_truth_results[['map']], ind = 2, shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 2), sub = 'Left - qtl2, Right - HaploQA') 
+png(file = file.path(root, 'do_qtl2_genoplot.png'), width = 1000, height = 800, res = 110)
+qtl2::plot_onegeno(do_results[['ph_geno']], do_results[['map']], ind = 1, shift = TRUE, main = paste0('Geno-wide genotypes from qtl2, individual ', 1), sub = 'Sample: DO')
+legend(13.5, 65, legend = c(names(qtl2::CCcolors)), title = 'Founder Colors', fill = qtl2::CCcolors)
+dev.off()
+
+
+png(file = file.path(root, 'cc_qtl2_genoplot.png'), width = 1000, height = 800, res = 110)
[…]

### 2022-11-28 app.R

@@ -2,9 +2,10 @@ library(rstudioapi)
 library(shiny)
 library(data.table)
 
-
 root <- dirname(getSourceEditorContext()$path)
-source(paste0(root,"/input_data_prep_functions.R"))
+
+source("input_data_prep_functions.R", local = T)
+#source(paste0(root,"/util.R"))
 
 ## results directory
 results_dir <- file.path(root, 'results')

### 2022-11-28 input_data_prep_functions.R

@@ -297,7 +297,7 @@ qtl2_ci <- function(summary_df, ngen, founder_haplo_lookup) {
 #
 # @return founders_total (data.frame) - combined dataframe with all founder individual files
 # columns of founders_total: sample_id, original_sample_id, snp_id, chromosome, position_bp, allele1_fwd, allele2_fwd, haplotype1, haplotype2
-get_founder_data <- function(founder_url, url_list, sample_type) {
+get_founder_data <- function(founder_url, url_list, sample_type, data_dir, founder_filename) {
   # get length to keep track of progress
   iter_len <- length(url_list)
   # loop through urls to extract individual files
@@ -305,10 +305,12 @@ get_founder_data <- function(founder_url, url_list, sample_type) {
   founders_sample_dir <- paste0(data_dir, '/', sample_type, '_founders/')
   dir.create(founders_sample_dir, showWarnings = FALSE) 
   
-  fp_founders <- paste0(root, '/', sample_type, '_founder_samples.csv')
+  fp_founders <- file.path(root, founder_filename)
   if (file.exists(fp_founders)) {
+    print('founder sample data exists')
     founders_total <- fread(fp_founders)
   } else { 
+    print('founder sample data does not exist')
     inc = 0
     for (url in url_list) {
       file <- sample_individual_scrape(url, url_domain) # call function from another script
@@ -368,7 +370,7 @@ get_haplotypes <- function(summary_df, data_dir) {
 # @return file_output (list) - nested list containing all files necessary for qtl2 input
 # specifically: df_geno, df_gmap, df_pmap, df_pheno, df_covar, df_foundergeno, df_crossinfo
 # see individual functions above for column names of each dataframe
-get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, summary_df, list_pheno, ngen, founders_list, marker_type, exclude_list, founder_url) {
+get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, summary_df, list_pheno, ngen, founders_list, marker_type, exclude_list, founder_url, founder_filename) {
   # list container to store outputs
   file_output <- list()
   
[…]

### 2022-12-07 draft_report.Rmd
Update draft_report.Rmd

@@
 In this project, we selected some GigaMUGA crosses, developed pipelines that retrieve genotyping data used in the initial HaploQA algorithm, performs haplotype reconstruction computations for both qtl2 genail and haploqa, and compare the percentages of markers within each method that agree with the optimal model of the according crosses to determine whether HaploQA or R/qtl2 gen-AIL would be the most appropriate for similar haplotype reconstruction tasks.
-
-The analysis was also done for MiniMUGA, as it is the standard cross type JAX uses for mutant mice, however, there is no tailored truth model for MiniMUGA.
+The analysis was also done for MiniMUGA, as it is the standard cross type JAX uses for mutant mice, however, there is no tailored truth model for MiniMUGA, so the amount of marker differences between HaploQA and qtl2 was used as the main metrics for MiniMUGA to determine the model performances.
 
@@
 
-The pipeline developed for this project is stored in the script, 'input_data_prep_functions.R', which is stored in the same working directory as this report and is sourced below:
+The main repository is https://github.com/TheJacksonLaboratory/HaploQA_qtl2_comp which contains the main pipelines developed for this project, all utility functions, and visualizations of selected results.
+
+The pipelines developed for this project is stored in the script, 'haplotype_reconstruction_pipelines.R', which is stored in the same working directory as this report and is sourced below:
 
@@
 As mentioned in the introduction, in order to decide whether to transition, the accuracy of haploqa and qtl2 each need to be compared to an optimal model.
-This project will perform haplotype reconstruction tasks using genortping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, <another one TBD>), 70 (?) individuals within the MiniMUGA sample.
+The pipelines in the sources script perform haplotype reconstruction tasks using genotyping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, and a GigaMUGA mutant cross), 70 individuals within the MiniMUGA sample.
 Decision was made to not include Muga/MegaMUGA data due to the discontinued usage of the array types.
 
-#### 2.1. Environment Set-up
+#### 2.2. Data Retrieval
 
-A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all the relevant file paths associated with each cross.
-The config file contains each individual crosstypes, their according marker annotation files, paths to the directory where the input data is stored and the directory where qtl2 input data is saved in, etc.
+Input data was retrieved from haploqa.jax.org, and converted into the qtl2 input format. The data wrangling consists of two parts:
+1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports for each individual within each cross
+2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input files for qtl2
 
+However, haploqa.jax.org contains a large number of uploaded samples and the input data for each cross needs to be retrieved from their according sample.
+For example, for cross 'Collaborative Cross', the URL to the sample is http://haploqa-dev.jax.org/tag/Collaborative%20Cross.html
+
+A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all relevant information associated with each cross such as the sample URL, input data directory, output data directory, marker annotation file names, etc.
 The purpose of this configuration file is to avoid having to pass a large list of parameters into each pipeline function.
@@
 ```{r}
-# comment it!
-#
+# cross type
 sample_type <- 'CC'
 ### Environment
-config_sample <- config[config$array_type == sample_type]
-marker_type <- config_sample$marker_type
-founders_list <- unlist(strsplit(config_sample$founders_list, ", "))
-n_founders <- length(founders_list)
-ngen <- config_sample$ngen
-sample_url <- config_sample$url
-founder_url <- config_sample$founder_url
+config_sample <- config[config$array_type == sample_type] # get config info for this cross type
+marker_type <- config_sample$marker_type # type of marker (GigaMUGA/MiniMUGA)
+founders_list <- unlist(strsplit(config_sample$founders_list, ", ")) # contributing founders
+n_founders <- length(founders_list) # number of founders used in reconstruction
+ngen <- config_sample$ngen # ngen parameter used in cross
+sample_url <- config_sample$url # URL to retrieve data from haploqa for this cross
+founder_url <- config_sample$founder_url # URL to retrieve founder data for GigaMUGA/MiniMUGA
+truth_model <- config_sample$truth_model # cross type of truth model for the cross
 
-# data output directory
+# directory to store input data retrieved from haploqa
 data_dir <- file.path(root, config_sample$data_dir)
+# directory to store input data for qtl2
+qtl2_dir <- file.path(root, config_sample$qtl2_dir)
 ```
 
-#### 2.2. Data Retrieval
-
-The pipeline takes the according data_dir, or qtl2_dir values for each cross, concatenate the variable values with the working directory (which is the directory where the functions are stored in), and read in/output files from the directories.
-
-Input data was retrieved from haploqa.jax.org, and converted into the qtl2 input format. The data wrangling consists of two parts:
-1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports for each individual within each cross
-2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input files for qtl2
-
-For example, for cross 'Collaborative Cross', the URL to the sample is http://haploqa-dev.jax.org/tag/Collaborative%20Cross.html
-
-The 'sample_summary_scrape' function reads the html elements of the website, (?)
-
-```{r, eval = FALSE} 
-## summary file
-haploqa_html <- read_html(sample_url)
-# extract information from html file
-html_temp <- haploqa_html %>% html_nodes("a") %>% html_attr("href")
-url_list <- paste0(url_domain, html_temp[grepl('/sample', html_temp)])
-summary_df <- sample_summary_scrape(haploqa_html, url_list, marker_type)
-
-## individual samples
-### this downloads the data directy to the folder, 'data_dir'. Try not to actually run this loop
-haploqa_html <- read_html(sample_url)
-# extract information from html file
-html_temp <- haploqa_html %>% html_nodes("a") %>% html_attr("href")
-url_list <- paste0(url_domain, html_temp[grepl('/sample', html_temp)])
-
-# list of urls to generate samples from
-url_ind <- unique(summary_df$`Sample Filepath`)
-
-# individual samples
-## 
-inc = 0
-for (url in url_ind) {
-  inc = inc + 1 # increment
-  file <- sample_individual_scrape(url, url_domain)
-  print(paste0('Working on file ', inc, '/', length(url_ind), ': ', file))
-  sample_df_save <- as.data.frame(content(GET(file)))
-  print('Writing to directory')
-  file_name <- unlist(strsplit(file, '/'))[6]
-  GET(file, write_disk(paste0(data_dir, '/', file_name), overwrite = TRUE), show_col_types = FALSE)
-}
-
-### note: get_qtl2_input reads all files from data_dir recursively and binds them into one big file.
-The big file is then merged with summary file to gather some necessary metadatas.
-This merged file is the main input file
-
-```
-
-The main input file, which contains (insert detailed info here later), is then processed into the required input format for qtl2.
-The qtl2 input file contains the following parts: genotype, genetic map, ... (do we need to explain thw input file structures to the audience or would they be familiar with this already?)
-
-There are three major pipelines:
+There are three major pipelines in the repository:
 a. 'haplotype_reconstruction_pipeline' - haplotype reconstructions for an entire cross (all individuals within the cross) 
@@
 
-whereas the cross, MiniMUGA, contains mutate mice where all individuals may have different contributing founder strains. Therefore, the MiniMUGA samples
+Some crosses, such as DO and CC, have the same contributing founders across all individuals.
+The haplotype reconstructions can therefore be conducted for all individuals in the entire cross at the same time.
+The pipeline 'haplotype_reconstruction_pipeline' can be used for crosses with consistent contributing founders for such purpose.
+Whereas the mutant crosses, such as MiniMUGA, contains individuals with mutate mice where all individuals may have different contributing founder strains.
+Haplotype reconstructions for these types of crosses need to conducted individually for each sample, and therefore, the pipeline 'sample_haplotype_reconstruction' should be used.
 
-Encoding(?) - in the main file, the retrieved data can be separated into two parts:
-    1. variables 'haplotype1', 'haplotype2', which represents the haploqa results,
-    2. variables 'allele1', 'allele2', which is the (original?) alleles used to calculate the haploqa results. These alleles will also be used for the qtl2 computations.
+The below results section will demonstrate how the pipelines are used for each cross and the results generated from them.
 
-Function 'get_qtl2_input' contains all the necessary algorithms to convert the data into separate required files. The output is then read in by qtl2 and used for calculations for different metrics.
-
-  1. qtl2 result calculations (genoprob(pr), phased geno(ph_geno), etc)
-  2. haploqa result calculations, with the same metrics as above.
-
-#### 2.3. Demo (?)
+#### 2.3. Results
 
 ##### 2.3.1. GigaMUGA
+
 The pipeline for GigaMUGA executes all the individuals within the sample. The contributing founders for all individuals within the GigaMUGA samples are always the same.
@@
 
-Results for DO:
+1. Results for DO:
+
 ```{r}
-# DO
 do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
@@
 Below are the genotype visualizations for some individuals:
-(pick some representative ones?)
+
 ```{r}
@@
 
-Results for CC:
+2. Results for CC:
+
 ```{r}
-# DO
 cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
@@
 
-genotype visualizations for some individuals: 
+Genotype visualizations for some individuals:
+
 ```{r}
@@
 
-##### 2.3.2. MiniMUGA
-MiniMUGA has a separate pipeline that runs for one individual at a time due to each individuals having different (contributing?) founders.
+3. Results for BXD F3:
 
 ```{r}
-sample1_results <- sample_haplotype_reconstruction('MiniMUGA', 'DJN', samples_gen = T, qtl2_file_gen = T)
-plot_onegeno_test(sample1_results[['ph_geno']], sample1_results[['ph_geno_haploqa']], sample1_results[['map']], ind = 1, 
+bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
+
+```
+
+Genotype visualizations for some individuals:
+
+```{r}
+plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 1, 
                           shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
@@
 ```{r}
-sample2_results <- sample_haplotype_reconstruction('MiniMUGA', 'JXN', samples_gen = T, qtl2_file_gen = T)
-plot_onegeno_test(sample2_results[['ph_geno']], sample2_results[['ph_geno_haploqa']], sample2_results[['map']], ind = 1, 
+plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 50, 
+                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
+                          sub = 'Left - qtl2, Right - HaploQA') 
+```
+
+4. Results for F2:
+
+```{r}
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
+```
+
+Genotype visualizations for some individuals:
+
+```{r}
+plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 1, 
                           shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                           sub = 'Left - qtl2, Right - HaploQA') 
+```
+
+```{r}
+plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 50, 
+                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
+                          sub = 'Left - qtl2, Right - HaploQA') 
+```
+
+5. GigaMUGA Mutant cross
+
+```{r}
+
+### GigaMUGA mutant
+gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
+url_domain <- 'http://haploqa-dev.jax.org/'
+marker_type <- 'GigaMUGA'
+html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
+url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
+summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
+summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
+summary_df <- summary_df[summary_df$`% No Call` < 10,]
+
+ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
+ind_gm <- ind_gm[!ind_gm %in% ind_gm[1:44]] # founder not in UWisc dictionary (SJL)
+ind_gm <- ind_gm[!ind_gm %in% ind_gm[4:24]] # empty - Q2-QP
+ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
+ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder
+ind_gm <- ind_gm[!ind_gm %in% c('U2', 'UA', 'UJ' ,'UT', 'W5', 'WD', 'WN', 'XZ', 'Y9', 'YH', 'YS')] # SJL 
+ind_gm <- ind_gm[!ind_gm %in% c('UK', 'UU', 'V4')] #C57BLKS/J
+ind_gm <- ind_gm[!ind_gm %in% c('WX', 'X7', 'X8', 'XF', 'XG', 'XQ', 'XR', 'XY', 'YG', 'YR')] # 129P3
+ind_gm <- ind_gm[!ind_gm %in% c('WY')] # FVB/NJ
+ind_gm <- ind_gm[!ind_gm %in% c('X2', 'X3', 'X4', 'X5', 'XA', 'XB', 'XC', 'XD', 'XJ', 'XK', 'XM', 'XN', 'XT', 'XU', 'XV', 'XW', 'Y2', 'Y3', 'Y4', 'Y5', 'YB', 'YC', 'YD', 'YE', 'YK', 'YM', 'YN', 'YP', 'YU', 'YV', 'YW', 'YX', 'Z5', 'Z6', 'Z7', 'ZC', 'ZD', 'ZE', 'ZF', 'ZM', 'ZN', 'ZP', 'ZQ', 'ZV', 'ZW', 'ZX', 'ZY', 'PY', 'Z3', 'ZB' ,'ZK', 'ZU')] # empty ones
+
+gigamuga_results <- list()
+for (ind in ind_gm) {
+  print(ind)
+  sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T)
+  gigamuga_results[[ind]] <- sample_gm_res
+}
+```
+
+##### 2.3.2. MiniMUGA
+
+```{r}
+### MiniMUGA pipeline
+minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
+url_domain <- 'http://haploqa-dev.jax.org/'
+marker_type <- 'MiniMUGA'
+html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
+url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
+summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
+ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
+ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
+ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
+ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
+minimuga_results <- list()
+mini_pct_res <- list()
+for (ind in ind_mini) {
+  #ind <- 'JXN'
+  print(ind)
+  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F)
+
+  df <- minimuga_geno_comp(sample_mini_res, ind, 'MiniMUGA')
+
+  minimuga_results[[ind]] <- sample_mini_res
+  mini_pct_res[[ind]] <- df
+}
+
 ```

### 2022-12-08 draft_report.Rmd
Update draft_report.Rmd

@@
 gigamuga_results <- list()
+gm_pct_res <- list()
 for (ind in ind_gm) {
@@
   sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T)
+  df <- minimuga_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
+
+  gm_pct_res[[ind]] <- df
   gigamuga_results[[ind]] <- sample_gm_res
@@
   sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F)
-
   df <- minimuga_geno_comp(sample_mini_res, ind, 'MiniMUGA')
@@
 ##### 3.2. Quantitative comparisons
-Combining the results of the truth model, the qtl2 haplot
-(Display the csvs, and perhaps the visualizations here too)
+
+For visualizations, we constructed boxplots that demonstrate the percentage of marker that do not have the same results between qtl2, haploQA and the truth model:
 
@@
 do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
-do_pct_df
-```
-
-```{r}
 # CC
 cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
-cc_pct_df
-#cc_pct_df
-#ggplot(cc_pct_df) + aes(x = qtl2_pct_diff, y = haploqa_pct_diff) + geom_point()
-```
-
-```{r}
 # BXD f3
 bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
-bxd_pct_df
-
-```
-
-```{r}
-# BXD f3
+# f2
 f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
-f2_pct_df
-
 ```
@@
 
-MiniMUGA
+The two mutant mice crosses, GigaMUGA Mutant and MiniMUGA, do not have a tailored optimal model. Therefore, plotting the percentage of marker discordance between qtl2 genail and haploQA.
+
+MiniMUGA:
 
@@
 
-#### 4. Conclusion - TBD
+GigaMUGA mutant:
+
+```{r}
+gm_pct_temp <- rbindlist(gm_pct_res)
+gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
+plot_df <- melt(gm_plot)
+plot_df$variable <- 'GigaMUGA Mutant'
+
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')
+
+```
+
+#### 4. Conclusion
 Overall, the existing results for qtl2 genail and haploqa show that the two methods have very similar performance for most of the crosses, with genail performing significantly better for the DO outbred mice samples.

### 2022-12-10 draft_report.Rmd
modularized scripts

@@
 root <- dirname(getSourceEditorContext()$path)
-source(paste0(root,"/input_data_prep_functions.R"))
+#source(paste0(root,"/input_data_prep_functions.R"))
 
-#source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
+source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
 
+# simulated phenotypes
+list_pheno <- c('WBC', 'NEUT')
 ```
@@
 
-```{r}
-# simulated phenotypes
-list_pheno <- c('WBC', 'NEUT')
-```
-
 1. Results for DO:
@@
 ```{r}
-do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
+do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
 
@@
 ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
-ind_gm <- ind_gm[!ind_gm %in% ind_gm[1:44]] # founder not in UWisc dictionary (SJL)
-ind_gm <- ind_gm[!ind_gm %in% ind_gm[4:24]] # empty - Q2-QP
+#ind_gm <- ind_gm[!ind_gm %in% ind_gm[1:44]] # founder not in UWisc dictionary (SJL)
+#ind_gm <- ind_gm[!ind_gm %in% ind_gm[4:24]] # empty - Q2-QP
 ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
 ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder
-ind_gm <- ind_gm[!ind_gm %in% c('U2', 'UA', 'UJ' ,'UT', 'W5', 'WD', 'WN', 'XZ', 'Y9', 'YH', 'YS')] # SJL 
-ind_gm <- ind_gm[!ind_gm %in% c('UK', 'UU', 'V4')] #C57BLKS/J
-ind_gm <- ind_gm[!ind_gm %in% c('WX', 'X7', 'X8', 'XF', 'XG', 'XQ', 'XR', 'XY', 'YG', 'YR')] # 129P3
-ind_gm <- ind_gm[!ind_gm %in% c('WY')] # FVB/NJ
-ind_gm <- ind_gm[!ind_gm %in% c('X2', 'X3', 'X4', 'X5', 'XA', 'XB', 'XC', 'XD', 'XJ', 'XK', 'XM', 'XN', 'XT', 'XU', 'XV', 'XW', 'Y2', 'Y3', 'Y4', 'Y5', 'YB', 'YC', 'YD', 'YE', 'YK', 'YM', 'YN', 'YP', 'YU', 'YV', 'YW', 'YX', 'Z5', 'Z6', 'Z7', 'ZC', 'ZD', 'ZE', 'ZF', 'ZM', 'ZN', 'ZP', 'ZQ', 'ZV', 'ZW', 'ZX', 'ZY', 'PY', 'Z3', 'ZB' ,'ZK', 'ZU')] # empty ones
+#ind_gm <- ind_gm[!ind_gm %in% c('U2', 'UA', 'UJ' ,'UT', 'W5', 'WD', 'WN', 'XZ', 'Y9', 'YH', 'YS')] # SJL 
+#ind_gm <- ind_gm[!ind_gm %in% c('UK', 'UU', 'V4')] #C57BLKS/J
+#ind_gm <- ind_gm[!ind_gm %in% c('WX', 'X7', 'X8', 'XF', 'XG', 'XQ', 'XR', 'XY', 'YG', 'YR')] # 129P3
+#ind_gm <- ind_gm[!ind_gm %in% c('WY')] # FVB/NJ
+#ind_gm <- ind_gm[!ind_gm %in% c('X2', 'X3', 'X4', 'X5', 'XA', 'XB', 'XC', 'XD', 'XJ', 'XK', 'XM', 'XN', 'XT', 'XU', 'XV', 'XW', 'Y2', 'Y3', 'Y4', 'Y5', 'YB', 'YC', 'YD', 'YE', 'YK', 'YM', 'YN', 'YP', 'YU', 'YV', 'YW', 'YX', 'Z5', 'Z6', 'Z7', 'ZC', 'ZD', 'ZE', 'ZF', 'ZM', 'ZN', 'ZP', 'ZQ', 'ZV', 'ZW', 'ZX', 'ZY', 'PY', 'Z3', 'ZB' ,'ZK', 'ZU')] # empty ones
 
@@
   sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T)
-  df <- minimuga_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
+  df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')
 
@@
   sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F)
-  df <- minimuga_geno_comp(sample_mini_res, ind, 'MiniMUGA')
+  df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')
 
@@
 
+mini_pct_res <- rbindlist(mini_pct_res)
 ```
@@
 
+(add screenshots of karyotype plots for outliers)
+
 The MiniMUGA samples are mutant mice samples, meaning they often have founders that contribute to only a few markers.
@@
 
+(make a plot to visualize run time)
+(talk about the strengths and weaknesses)
+(figure out the error LOD)
+qtl2 does not scale well with large number of founders
+
 #### 5. References
@@
 PMID: 30591514; PMCID: PMC6366910.
+
+qtl2fst - store large objects 
+calc_genoprob_fst
+(allele prob fst)
+for DO - 
+time the regular genoprob and fst
+get the file size/memory usage of RDS and FST

### 2022-12-10 get_comp_results.R

@@ -1,0 +1,536 @@
+## contains all functions to calculate accuracy metrics and/or to compare results 
+### function directory
+# 1. location_xo_comp - calculates the distances between crossover locations of haploqa and qtl2 results
+# 2. err_comp
+# 3. comp_df_int
+
+library(rstudioapi)
+root <- dirname(getSourceEditorContext()$path)
+
+# calculates the distance between crossover locations for the two models
+# 1. compares the location crossovers values (which are usually lists) between haploqa and qtl2
+# 2. determine which one has more crossovers than the other (length of lists)
+# 3. maps the values in the shorter crossover list each to the closest value in the longer list
+# 4. finds the absolute value between the pair
+#
+# @param qtl2_pos (listed matrices) - output of locate_xo() for qtl2
+# @param haploqa_pos (listed matrices) - output of locate_xo() for haploqa
+# @param num_chr (list) - list of chromosomes
+#
+# @output x_loc_diff (listed dataframe) - results of the calculations, with the results for each chromosome being an element of the list
+location_xo_comp <- function(qtl2_pos, haploqa_pos, num_chr) {
+  x_loc_diff <- list()
+  for (chr in num_chr) {
+    print(chr)
+    pos_h_chr <- haploqa_pos[[chr]]
+    pos_chr <- qtl2_pos[[chr]]
+    nmar_h <- sapply(pos_h_chr, length)
+    nmar <- sapply(pos_chr, length) # none has the same length...
+    #x_loc_diff[[chr]] <- list()
+    for (i in seq(1:length(pos_chr))) {
+      #i <- 149
+      x <- unlist(pos_h_chr[i])
+      y <- unlist(pos_chr[i])
+      if (length(x) > length(y)) {
+        df <- apply(abs(outer(x, y, '-')), 2, min)
+      }
+      if (length(x) < length(y)) {
+        df <- apply(abs(outer(y, x, '-')), 2, min)
+      }
+      x_loc_diff[[chr]][[i]] <- ifelse(is.na(mean(df)), 0, mean(df))
+    } # make a histogram of some sort, 100 bins
+  }
+  return(x_loc_diff)
+}
+
+
[…]

### 2022-12-10 haplotype_reconstruction_pipelines.R

@@ -1,0 +1,803 @@
+### Contains three major pipelines for haplotype reconstructions
+
+library(rstudioapi)
+root <- dirname(getSourceEditorContext()$path)
+
+# environment set-up
+source(paste0(root,"/input_data_retrieval.R"))
+source(paste0(root,"/qtl2_input_data_prep.R"))
+source(paste0(root,"/qtl2_sim_comp.R"))
+source(paste0(root,"/get_comp_results.R"))
+source(paste0(root,"/utils.R"))
+
+
+# function to initiate the haplotype reconstruction pipeline
+# @param sample_type (string) - type of sample, such as CC, DO, F2, etc.
+# @param list_pheno (list) - list of phenotypes to be simulated
+# @param qtl2_file_gen (True/False) - toggle to set whether to (re)generate qtl2 input files
+# @param samples_gen (True/False) - toggle to set whether to (re)generate individual sample  files
+#
+# @return results (list of dataframes) - a list containing all objects necessary for computations
+haplotype_reconstruction_pipeline <- function(sample_type, list_pheno, qtl2_file_gen, samples_gen) {
+  
+  #sample_type <- 'CC'
+  #truth_model = T
+  ## results directory
+  results_dir <- file.path(root, 'results')
+  dir.create(results_dir, showWarnings = FALSE) 
+  
+  num_chr <- c((1:19),"X")
+  
+  ### config file
+  config <- fread(paste0(root, '/annotations_config.csv'))
+  
+  ### Environment
+  config_sample <- config[config$array_type == sample_type]
+  marker_type <- config_sample$marker_type
+  founders_list <- unlist(strsplit(config_sample$founders_list, ", "))
+  n_founders <- length(founders_list)
+  ngen <- config_sample$ngen
+  sample_url <- config_sample$url
+  founder_url <- config_sample$founder_url
+  founder_filename <- config_sample$founder_filename
+  
+  # data output directory
+  data_dir <- file.path(root, config_sample$data_dir)
+  dir.create(data_dir, showWarnings = FALSE) 
+  
+  ## create a data directory for qtl2 input data
+  qtl2_dir <- file.path(root, config_sample$qtl2_dir) # name of desired output folder
+  # create if folder not exist
+  dir.create(qtl2_dir, showWarnings = FALSE)
+  
+  # filepaths to save any rds file
[…]

### 2022-12-10 input_data_retrieval.R

@@ -1,0 +1,71 @@
+### contains all functions necessary to retrieve initial input files from haploqoa
+### function directory
+## 1. sample_summary_scrape - get metadata table for samples from haploqa website
+## 2. sample_individual_scrape - get haplotype calls/input individual files 
+## 3. read_sample_txt - function to read txt and perform some pre-processing
+
+library(rstudioapi)
+root <- dirname(getSourceEditorContext()$path)
+
+
+# function to retrieve summary table from haploqa website
+# basically takes the table shown on main page as-is
+# @param html_file (html_document) - html of main page sample website on HaploQA, output of read_html(url)
+# @param url_list (list) - url to main page of individuals in said sample, bound on summary table as metadata
+#
+# @return sum_table (data.frame) - table as shown on HaploQA sample website
+# columns of sum_table: ID (sample IDs), secondary ID, Haplotype Candidate (T/F), Strain Name, Sex, % Het Calls, % Hom Calls, % No Call, % Concordance, Sample filepath
+sample_summary_scrape <- function(html_file, url_list, marker_type) {
+  # convert into table
+  html_table <- html_file %>% html_nodes(".table") %>% html_table()
+  sum_table <- html_table[[1]][-1] # remove first column, as it's blank
+  sum_table <- as.data.frame(sum_table, row.names = F) # convert into dataframe
+  # clean up ID columns
+  sum_table$`ID (Secondary IDs)` <- gsub("\\s+", "", sum_table$`ID (Secondary IDs)`) # remove spaces
+  # split IDs and secondary IDs
+  sum_table <- sum_table %>%
+    separate(`ID (Secondary IDs)`, c("ID", "Secondary IDs"), "\\(")
+  sum_table$`Secondary IDs` <- gsub("\\)", "", sum_table$`Secondary IDs`) #remove parentheses
+  sum_table$`Sample Filepath` <- url_list
+  if (!missing(marker_type)) {
+    sum_table <- sum_table[sum_table$Platform == marker_type,]
+  }
+  return(sum_table)
+}
+
+# function to retrieve individuals in a given HaploQA sample
+# @param url (string) - url of main page of each individual 
+# @param url_domain (string) - almost always 'http://haploqa-dev.jax.org/' 
+#
+# @return file (string) - url to direct download of the file associated with individual
+sample_individual_scrape <- function(url, url_domain) {
+  # read html of website with sample data
+  html_sample <- read_html(url)
[…]

### 2022-12-10 qtl2_input_data_prep.R

@@ -1,0 +1,406 @@
+### Contains all functions to generate qtl2 input data
+
+library(rstudioapi)
+root <- dirname(getSourceEditorContext()$path)
+
+source(paste0(root,"/input_data_retrieval.R"))
+source(paste0(root,"/utils.R"))
+
+
+# function to generate genotype data for qtl2 input
+# @param df (data.frame) - combined dataframe containing all individual files with only necessary columns selected, merged with summary table and chr 0,Y,M and no calls>10% removed
+# columns of df: sample_id, original_sample_id, snp_id, chromosome, position_bp, allele1_fwd, allele2_fwd, haplotype1, haplotype2, Sex, % Het. Calls, % No Call, Platform
+# @param annot_encode_df (data.frame) - processed UWisc annotation file with alleles
+# columns of annot_encode_df: marker, chr, allele1, allele2
+#
+# @return df_geno_encoded (data.frame) - converted genotype data in qtl2 required format
+qtl2_geno <- function(df, annot_encode_df) {
+  ### genotype data
+  geno_sub <- df %>% select(sample_id, snp_id, allele1_fwd, allele2_fwd) %>%
+    mutate(gene_exp = paste(allele1_fwd, allele2_fwd, sep = '')) %>%
+    select(sample_id, snp_id, gene_exp) %>% rename(marker = snp_id)
+  df_geno <- dcast(geno_sub, marker~sample_id, value.var="gene_exp") # inbred - most are homozygous
+  ### cut markers more than 10%
+  # merge with annotations
+  df_geno <- df_geno %>% filter(marker %in% unique(annot_encode_df$marker)) ### replace with a stopif assert line here
+  # put alleles in order of marker as shown in geno dataframe
+  geno_encode_annot <- merge(df_geno, annot_encode_df, by = 'marker', all.x = T) %>% select(allele1, allele2) # let it break if there's NAs
+  rownames(df_geno) <- df_geno$marker
+  df_geno_encoded <- as.data.frame(encode_geno(df_geno[,-1], geno_encode_annot))
+  df_geno_encoded$marker <- rownames(df_geno)
+  # reorder columns
+  df_geno_encoded <- df_geno_encoded %>% select(marker, everything())
+  colnames(df_geno_encoded) <- colnames(df_geno)
+  return(df_geno_encoded)
+}
+
+# function to generate genetic and physical map data for qtl2 input
[…]

### 2022-12-10 qtl2_sim_comp.R

@@ -1,0 +1,227 @@
+# contains all functions that aligns haploqa results with that of qtl2 and save the results
+# this is necessary because the computed haploqa results needs to be in the same format as qtl2 (with genoprob, phased geno, etc.)
+### function directory
+## 1. maxmarg_sim - simulate maxmarg phasing computations for haploqa
+## 2. save_rds - save all results into rds
+
+
+library(rstudioapi)
+root <- dirname(getSourceEditorContext()$path)
+
+
+source(paste0(root,"/utils.R"))
+
+# function to simulate the qtl2 maxmarg phasing computations
+# @param summary_df (data.frame) - summary table scraped from haploqa
+# @param data_dir (string) - directory where raw input data is stored
+# @param num_chr (vector) - chromosomes present, sorted into order
+# @param founder_all_rev_lookup (named list) - lookup table to convert number codes (1,2,3,4) into letters(AA, BB, CC, DD)
+# @param cross (cross object) - output of read_cross for the sample model
+# @param qtl2_dir (string) - directory where qtl2 input data is stored
+# @param sample_type (string) - type of sample, such as CC, DO, F2, etc.
+# @param n_founders (int) - number of founders exist in the model
+# @param founder_haplo_lookup (named list) - lookup table to convert strain names (A/J, C57BL/6J) into codes (A, B)
+#
+# @return ginf_haploqa (list of matrices) - phased genotypes indexed by chromosome, same as output of maxmarg() from qtl2
+maxmarg_sim <- function(summary_df, data_dir, num_chr, founder_all_rev_lookup, cross, qtl2_dir, sample_type, n_founders, founder_haplo_lookup, pr) {
+  
+  df_cov <- fread(file.path(qtl2_dir, 'test_covar.csv')) %>% rename('sample_id' = 'id')
+  
+  df_all <- get_haplotypes(summary_df, data_dir)
+  df_test <- df_all %>% select(sample_id, snp_id, haplotype1, haplotype2, chromosome) %>% unique()
+  df_test[,c(3,4)] <- as.data.frame(apply(df_test[,c(3,4)], 2, function(x) founder_haplo_lookup[x]))
+  
+  # gender mapping
+  df_test <- merge(df_test, df_cov, by = 'sample_id')
[…]

### 2022-12-10 Scripts/utils.R
Modularizing scripts

@@ -1,0 +1,646 @@
+### Contains all general utility functions used across the board
+
+
+
+# function to retrieve only the dataframe that all qtl2 input dataframes were based on for testing purposes, and/or the haplotype columns
+# @param summary_df (data.frame) - table as shown on HaploQA sample website, same as output of sample_summary_scrape
+# @param data_dir (string) - filepath to where individual files (outputs of sample_individual_scrape) are stored
+#
+# @return df_all (data.frame) - dataframe with haplotype columns
+get_haplotypes <- function(summary_df, data_dir) {
+  # all txt file from directory
+  data_files <- dir(data_dir, pattern = '\\.txt$', full.names = TRUE)
+  ### combine all files
+  df_raw <- rbindlist(lapply(data_files, read_sample_txt)) # save Y chrom here for gender plotting
+  
+  # clean up summary df
+  sum_df <- summary_df %>% 
+    select(ID, Sex, `% Het. Calls`, `% No Call`, Platform) %>% rename(sample_id = ID)
+  sum_df$`% Het. Calls` <- as.numeric(gsub("%", "", sum_df$`% Het. Calls`))
+  sum_df$`% No Call` <- as.numeric(gsub("%", "", sum_df$`% No Call`))
+  
+  # eliminate those that has no call > 10%
+  ### use checkifnot to make sure the markers/SNP are in the annotation file (annotation file is the boss)
+  df_all <- merge(df_raw, sum_df, by = 'sample_id') %>% filter(`% No Call` < 10) %>% filter(!chromosome %in% c('0', 'Y', 'M'))
+  
+  return(df_all)
+}
+
+# function to sort chromosomes into custom order
+# @param df (data.frame) - selected dataframe to be sorted
+# @param sort_order (list) - order to sort the chromosomes
+#
+# @return df (data.frame) - dataframe with chromosomes sorted in the selected order
+sort_chr <- function(df, sort_order) {
+  df$chr <- factor(df$chr, levels = sort_order)
+  df <- df[order(df$chr),] 
+  return(df)
+}
+
+
+# function to add cross attributes to an object
+# @param crosstype (string) - type of cross model
+# @param num_chr (string) - chromosome to be mapped
+# @param chromosomes (string) - chromosomes present in the model
[…]

### 2022-12-10 input_data_retrieval.R
Delete input_data_retrieval.R

@@ -1,71 +1,0 @@
-### contains all functions necessary to retrieve initial input files from haploqoa
-### function directory
-## 1. sample_summary_scrape - get metadata table for samples from haploqa website
-## 2. sample_individual_scrape - get haplotype calls/input individual files 
-## 3. read_sample_txt - function to read txt and perform some pre-processing
-
-library(rstudioapi)
-root <- dirname(getSourceEditorContext()$path)
-
-
-# function to retrieve summary table from haploqa website
-# basically takes the table shown on main page as-is
-# @param html_file (html_document) - html of main page sample website on HaploQA, output of read_html(url)
-# @param url_list (list) - url to main page of individuals in said sample, bound on summary table as metadata
-#
-# @return sum_table (data.frame) - table as shown on HaploQA sample website
-# columns of sum_table: ID (sample IDs), secondary ID, Haplotype Candidate (T/F), Strain Name, Sex, % Het Calls, % Hom Calls, % No Call, % Concordance, Sample filepath
-sample_summary_scrape <- function(html_file, url_list, marker_type) {
-  # convert into table
-  html_table <- html_file %>% html_nodes(".table") %>% html_table()
-  sum_table <- html_table[[1]][-1] # remove first column, as it's blank
-  sum_table <- as.data.frame(sum_table, row.names = F) # convert into dataframe
-  # clean up ID columns
-  sum_table$`ID (Secondary IDs)` <- gsub("\\s+", "", sum_table$`ID (Secondary IDs)`) # remove spaces
-  # split IDs and secondary IDs
-  sum_table <- sum_table %>%
-    separate(`ID (Secondary IDs)`, c("ID", "Secondary IDs"), "\\(")
-  sum_table$`Secondary IDs` <- gsub("\\)", "", sum_table$`Secondary IDs`) #remove parentheses
-  sum_table$`Sample Filepath` <- url_list
-  if (!missing(marker_type)) {
-    sum_table <- sum_table[sum_table$Platform == marker_type,]
-  }
-  return(sum_table)
-}
-
-# function to retrieve individuals in a given HaploQA sample
-# @param url (string) - url of main page of each individual 
-# @param url_domain (string) - almost always 'http://haploqa-dev.jax.org/' 
-#
-# @return file (string) - url to direct download of the file associated with individual
-sample_individual_scrape <- function(url, url_domain) {
-  # read html of website with sample data
-  html_sample <- read_html(url)
[…]

### 2022-12-11 Scripts/qtl2_input_data_prep.R
fixed errors

@@ -3,8 +3,8 @@
 library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
 
-source(paste0(root,"/input_data_retrieval.R"))
-source(paste0(root,"/utils.R"))
+source(paste0(root,"/Scripts/input_data_retrieval.R"))
+source(paste0(root,"/Scripts/utils.R"))
 
 
 # function to generate genotype data for qtl2 input

### 2022-12-11 Scripts/qtl2_sim_comp.R

@@ -9,7 +9,7 @@ library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
 
 
-source(paste0(root,"/utils.R"))
+source(paste0(root,"/Scripts/utils.R"))
 
 # function to simulate the qtl2 maxmarg phasing computations
 # @param summary_df (data.frame) - summary table scraped from haploqa
@@ -101,7 +101,7 @@ maxmarg_sim <- function(summary_df, data_dir, num_chr, founder_all_rev_lookup, c
 #
 # @return cross_list (list containing multiple objects) - a list that has all the objects necessary for further research
 # such as genetic map, calculated qtl2 objects, phased genotypes for both qtl2 and haploqa, etc.
-save_rds <- function(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_founders, founder_haplo_lookup, truth_model) {
+qtl2_metrics_comp <- function(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_founders, founder_haplo_lookup, truth_model) {
   num_chr <- c((1:19),"X")
   print(sample_type)
   if (truth_model == T) {
@@ -190,6 +190,8 @@ save_rds <- function(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_fo
   
   cross_list[['data_dir']] <- data_dir
   
+  cross_list[['founder_lookup']] <- founder_haplo_lookup
+  
   founder_lookup_table <- fread(file.path(root, 'founder_lookup_table.csv'))
   founder_all_rev_lookup <- setNames(founder_lookup_table$founder_codes, founder_lookup_table$founder_id)
   
@@ -198,7 +200,7 @@ save_rds <- function(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_fo
     ginf_haploqa_fp <- paste0(rds_dir, "/ginf_haploqa_", sample_type, ".rds")
     if (!file.exists(ginf_haploqa_fp)) {
       print('haploqa ginf file does not exist, running calculations from maxmarg simulation')
-      ginf_haploqa <- maxmarg_sim(summary_df, data_dir, num_chr, founder_all_rev_lookup, cross = cross_list[['cross']], qtl2_dir, sample_type, n_founders, founder_haplo_lookup)
[…]

### 2022-12-13 Scripts/qtl2_input_data_prep.R
12/12 update

@@ -254,15 +254,28 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
   } else { 
     print('founder sample data does not exist')
     inc = 0
+    st_error_urls <- list()
     for (url in url_list) {
       file <- sample_individual_scrape(url, url_domain) # call function from another script
       inc = inc+1 # track progress
       print(paste0('working on file ', inc, '/', iter_len, ' ', file))
-      df_temp <- as.data.frame(content(GET(file), encoding="UTF-8"))
+      # screen time errors
+      skip <- FALSE
+      tryCatch(as.data.frame(content(GET(file), encoding="UTF-8")), error = function(e) { skip <<- TRUE})
+      if(!skip) { 
+        df_temp <- as.data.frame(content(GET(file), encoding="UTF-8"))
+      } else {
+        print(paste0('skipped ', url, ' due to screentime error'))
+        st_error_urls <- c(st_error_urls, url)
+        next
+      } 
+      
       file_name <- unlist(strsplit(file, '/'))[6]
       print('Writing to directory')
       #founders_total <- rbind(founders_total,df_temp)
+      
       GET(file, write_disk(file.path(founders_sample_dir, file_name), overwrite = TRUE), show_col_types = FALSE)
+      
     }
     
     founder_data_files <- dir(founders_sample_dir, pattern = '\\.txt$', full.names = TRUE)
@@ -271,6 +284,7 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
     write.csv(founders_total, fp_founders, row.names = F)
   }
   
+  print(paste0('urls skipped: ', st_error_urls))
   
   return(founders_total)
   
@@ -302,7 +316,7 @@ get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, s
   # all txt file from directory
   data_files <- dir(data_dir, pattern = '\\.txt$', full.names = TRUE)
   ### combine all files
-  df_raw <- rbindlist(lapply(data_files, read_sample_txt)) %>% filter(!sample_id %in% exclude_list) # save Y chrom here for gender plotting
+  df_raw <- rbindlist(lapply(data_files, read_sample_txt)) #%>% filter(!sample_id %in% exclude_list) # save Y chrom here for gender plotting
   
   ## exclude the ones with bad string names
   if(sample_type == 'CC') {
[…]

### 2022-12-13 Scripts/qtl2_sim_comp.R

@@ -87,10 +87,6 @@ maxmarg_sim <- function(summary_df, data_dir, num_chr, founder_all_rev_lookup, c
   return(ginf_haploqa)
 }
 
-
-
-
-
 # function to store results into rds files for future use
 # @param rds_dir (string) - directory for which the rds files should be saved in
 # @param qtl2_dir (string) - directory to read the qtl2 input files from

### 2022-12-13 haplotype_reconstruction_pipelines.R

@@ -4,11 +4,11 @@ library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
 
 # environment set-up
-source(paste0(root,"/input_data_retrieval.R"))
-source(paste0(root,"/qtl2_input_data_prep.R"))
-source(paste0(root,"/qtl2_sim_comp.R"))
-source(paste0(root,"/get_comp_results.R"))
-source(paste0(root,"/utils.R"))
+source(paste0(root,"/Scripts/input_data_retrieval.R"))
+source(paste0(root,"/Scripts/qtl2_input_data_prep.R"))
+source(paste0(root,"/Scripts/qtl2_sim_comp.R"))
+source(paste0(root,"/Scripts/get_comp_results.R"))
+source(paste0(root,"/Scripts/utils.R"))
 
 
 # function to initiate the haplotype reconstruction pipeline

### 2022-12-14 draft_report.Rmd
12/14 updates

@@
 According to the HaploQA about website, 'On the import of a sample, the haplotype probabilities that are represented in the plots are calculated.' and the haplotype reconstruction process for the contributing founder strains is then done based on the calculated probabilities.
-The source codes of HaploQA was written in Python 3.
+The source codes of HaploQA was written in Python 3 and JavaScript.
 
@@
 Some crosses, such as DO and CC, have the same contributing founders across all individuals.
-The haplotype reconstructions can therefore be conducted for all individuals in the entire cross at the same time.
-The pipeline 'haplotype_reconstruction_pipeline' can be used for crosses with consistent contributing founders for such purpose.
+The haplotype reconstructions can therefore be conducted for all individuals in the entire cross at the same time, and the pipeline 'haplotype_reconstruction_pipeline' can be used for crosses with consistent contributing founders for such purpose.
 Whereas the mutant crosses, such as MiniMUGA, contains individuals with mutate mice where all individuals may have different contributing founder strains.
@@
 
-```{r}
+```{r message=FALSE}
 do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
@@
 ```{r}
-cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F, truth_model = F)
+cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
 

### 2022-12-14 Scripts/qtl2_input_data_prep.R

@@ -240,6 +240,7 @@ qtl2_ci <- function(summary_df, ngen, founder_haplo_lookup) {
 # @return founders_total (data.frame) - combined dataframe with all founder individual files
 # columns of founders_total: sample_id, original_sample_id, snp_id, chromosome, position_bp, allele1_fwd, allele2_fwd, haplotype1, haplotype2
 get_founder_data <- function(founder_url, url_list, sample_type, data_dir, founder_filename) {
+  
   # get length to keep track of progress
   iter_len <- length(url_list)
   # loop through urls to extract individual files
@@ -261,7 +262,7 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
       print(paste0('working on file ', inc, '/', iter_len, ' ', file))
       # screen time errors
       skip <- FALSE
-      tryCatch(as.data.frame(content(GET(file), encoding="UTF-8")), error = function(e) { skip <<- TRUE})
+      tryCatch(as.data.frame(content(GET(file), encoding="UTF-8")), error = function(e) {skip <<- TRUE})
       if(!skip) { 
         df_temp <- as.data.frame(content(GET(file), encoding="UTF-8"))
       } else {
@@ -284,7 +285,10 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
     write.csv(founders_total, fp_founders, row.names = F)
   }
   
-  print(paste0('urls skipped: ', st_error_urls))
+  ## comments, make it into a log file
+  sum_temp <- sample_summary_scrape(read_html(founder_url), url_list, marker_type)
+  skipped_ids <- sum_temp[sum_temp$`Sample Filepath` %in% st_error_urls,]$ID
+  print(paste0('urls skipped: ', skipped_ids))
   
   return(founders_total)
   

### 2022-12-14 Scripts/qtl2_sim_comp.R

@@ -58,7 +58,7 @@ maxmarg_sim <- function(summary_df, data_dir, num_chr, founder_all_rev_lookup, c
   ## save per chromosome
   ginf_haploqa <- list()
   for (i in num_chr) {
-     i <- '1'
+     #i <- '1'
     
     # founder table
     geno_codes <- colnames(pr[[i]])

### 2022-12-14 haplotype_reconstruction_pipelines.R

@@ -139,17 +139,28 @@ haplotype_reconstruction_pipeline <- function(sample_type, list_pheno, qtl2_file
     
     # individual samples
     inc = 0
+    st_error_urls <- list()
     for (url in url_ind) {
       inc = inc + 1 # increment
-      file <- sample_individual_scrape(url, url_domain)
+      # screen time errors
+      skip <- FALSE
       print(paste0('Working on file ', inc, '/', length(url_ind), ': ', file))
-      sample_df_save <- as.data.frame(content(GET(file)))
+      file <- sample_individual_scrape(url, url_domain)
+      tryCatch(as.data.frame(content(GET(file), encoding="UTF-8")), error = function(e) {skip <<- TRUE})
+      if(!skip) { 
+        sample_df_save <- as.data.frame(content(GET(file)))
+      } else {
+        print(paste0('skipped ', url, ' due to screentime error'))
+        st_error_urls <- c(st_error_urls, url)
+        next
+      } 
       print('Writing to directory')
       file_name <- unlist(strsplit(file, '/'))[6]
       GET(file, write_disk(paste0(data_dir, '/', file_name), overwrite = TRUE), show_col_types = FALSE)
       
       
     }
+    print(paste0('urls skipped: ', st_error_urls))
   }
   
   ### Part 2 - convert results to qtl2 input format

### 2022-12-16 draft_report.Rmd
updated scripts

@@
 source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
+results_dir <- file.path(root, 'results')
 
@@
 
-```{r}
+```{r message=FALSE}
 cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
@@
 ```{r}
-bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
+bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)
 
@@
 ```{r}
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = F)
 ```
@@
 ```{r}
-
 ### GigaMUGA mutant
@@
   print(ind)
-  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F)
+  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
   df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')
@@
 library(ggplot2)    
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('qtl2 - proportion concordance between optimal and implemented models')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('qtl2 - proportion discordance between optimal and implemented models')
 ```
 
-According to the visualization, which shows the percentage of marker differences between qtl2 genail and the truth models - DO, F3 and F2 are all producing results that are almost exactly the same as the optimal model, with CC having some discrepancies.
+According to the visualization, which shows the percentage of marker discordance between qtl2 genail and the truth models - DO, F3 and F2 are all producing results that are almost exactly the same as the optimal model, with CC having some discrepancies.
 We believe the reason why CC performs so differently is that the optimal model for CC does not allow any heterozygosity, whereas the genail does, so the genail model showed some residual heterozygosity from the CC mice which the optimal model does not have the ability to show.
+
+The outliers at 80% discordance are samples V6 and VP.
+Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate.
+However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous.
+Further investigation may be needed on the two samples.
+
+Below are the karyotype plots for the two samples:
+
+![Karyotype plot - sample VP in CC](/Users/linc/Desktop/VP_geno_cc.png) ![Karyotype plot - sample V6 in CC](/Users/linc/Desktop/V6_geno_cc.png)
 
@@
 library(ggplot2)    
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa - proportion concordance between optimal and implemented models')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa - proportion discordance between optimal and implemented models')
 ```
@@
 library(ggplot2)    
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('proportion concordance between implemented models (haploqa vs qtl2)')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa vs qtl2 - proportion discordance between implemented models')
 ```
@@
 
-(add screenshots of karyotype plots for outliers)
-
 The MiniMUGA samples are mutant mice samples, meaning they often have founders that contribute to only a few markers.
@@
 
-(make a plot to visualize run time)
-(talk about the strengths and weaknesses)
-(figure out the error LOD)
-qtl2 does not scale well with large number of founders
+Aside from the statistical metrics comparison between the haplotype reconstruction results, it is also worth taking the programmatic strengths and weaknesses of the two models into consideration:
+
+HaploQA:
+  1. advantages: 
+      a. Not a programming tool - people without relevant knowledge can still use the tool and perform haplotype reconstruction
+      b. Works for all MUGA array-based genotyping data 
+      c. Detailed and concise visualizations on web interface
+  2. disadvantages:
+      a. haplotype reports are not easy to access computationally
+      b. source codes are also not easy to access computationally
+      c. expensive to maintain at the organization's costs
+
+R/qtl2: 
+  1. advantages:
+      a. fully accessible computationally
+      b. capable of performing other genetic analysis
+      c. currently maintained by developers
+  2. disadvantages:
+      a. algorithms is probability based, tends to bounce states when low-contribution founders are present
+      b. does not scale well with large number of founders, run time increases exponentially 
+      c. visualizations are less detailed and need to be generated on demand
 
@@
 PMID: 30591514; PMCID: PMC6366910.
-
-qtl2fst - store large objects 
-calc_genoprob_fst
-(allele prob fst)
-for DO - 
-time the regular genoprob and fst
-get the file size/memory usage of RDS and FST

### 2022-12-16 Scripts/qtl2_input_data_prep.R

@@ -279,17 +279,18 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
       
     }
     
+    ## comments, make it into a log file
+    sum_temp <- sample_summary_scrape(read_html(founder_url), url_list, marker_type)
+    skipped_ids <- sum_temp[sum_temp$`Sample Filepath` %in% st_error_urls,]$ID
+    print(paste0('urls skipped: ', skipped_ids))
+    
     founder_data_files <- dir(founders_sample_dir, pattern = '\\.txt$', full.names = TRUE)
     ### combine all files
     founders_total <- rbindlist(lapply(founder_data_files, read_sample_txt))
     write.csv(founders_total, fp_founders, row.names = F)
   }
   
-  ## comments, make it into a log file
-  sum_temp <- sample_summary_scrape(read_html(founder_url), url_list, marker_type)
-  skipped_ids <- sum_temp[sum_temp$`Sample Filepath` %in% st_error_urls,]$ID
-  print(paste0('urls skipped: ', skipped_ids))
-  
+
   return(founders_total)
   
 }

### 2022-12-16 haplotype_reconstruction_pipelines.R

@@ -390,7 +390,7 @@ sample_haplotype_reconstruction <- function(sample_type, sample_name, samples_ge
     founder_sum <- founder_sum[founder_sum$`Haplotype Candidate` == 'True',] %>% rename(sample_id = ID, original_sample_id = 'Secondary IDs')
     fp_founders <- file.path(root, 'MiniMUGA_founder_samples.csv')
     
-    founders_total <- get_founder_data(sample_url, founder_sum$`Sample Filepath`, sample_type, data_dir, founder_filename)
+    founders_total <- get_founder_data(sample_url, founder_sum$`Sample Filepath`, sample_type, data_dir, 'MiniMUGA_founders.csv')
     
     founders_list <- names(founder_haplo_lookup)
     

### 2022-12-16 draft_report.Rmd
updated details

@@
 results_dir <- file.path(root, 'results')
+shiny_pct_fp <- file.path(results_dir, 'shiny_pct_csvs')
 
@@
 ```{r}
-#root <- dirname(rstudioapi::getSourceEditorContext()$path)
 config <- fread(paste0(root, '/annotations_config.csv'))
@@
 ```{r}
-bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)
+bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
 
@@
 ```{r}
-plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 50, 
+plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 5, 
                           shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
@@
 ```{r}
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = F)
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
 ```
@@
 ```{r}
-plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 50, 
+plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 5, 
                           shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
@@
 ```{r}
-### GigaMUGA mutant
+# ### GigaMUGA mutant
 gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
@@
 ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
-#ind_gm <- ind_gm[!ind_gm %in% ind_gm[1:44]] # founder not in UWisc dictionary (SJL)
-#ind_gm <- ind_gm[!ind_gm %in% ind_gm[4:24]] # empty - Q2-QP
 ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
 ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder
-#ind_gm <- ind_gm[!ind_gm %in% c('U2', 'UA', 'UJ' ,'UT', 'W5', 'WD', 'WN', 'XZ', 'Y9', 'YH', 'YS')] # SJL 
-#ind_gm <- ind_gm[!ind_gm %in% c('UK', 'UU', 'V4')] #C57BLKS/J
-#ind_gm <- ind_gm[!ind_gm %in% c('WX', 'X7', 'X8', 'XF', 'XG', 'XQ', 'XR', 'XY', 'YG', 'YR')] # 129P3
-#ind_gm <- ind_gm[!ind_gm %in% c('WY')] # FVB/NJ
-#ind_gm <- ind_gm[!ind_gm %in% c('X2', 'X3', 'X4', 'X5', 'XA', 'XB', 'XC', 'XD', 'XJ', 'XK', 'XM', 'XN', 'XT', 'XU', 'XV', 'XW', 'Y2', 'Y3', 'Y4', 'Y5', 'YB', 'YC', 'YD', 'YE', 'YK', 'YM', 'YN', 'YP', 'YU', 'YV', 'YW', 'YX', 'Z5', 'Z6', 'Z7', 'ZC', 'ZD', 'ZE', 'ZF', 'ZM', 'ZN', 'ZP', 'ZQ', 'ZV', 'ZW', 'ZX', 'ZY', 'PY', 'Z3', 'ZB' ,'ZK', 'ZU')] # empty ones
 
@@
 ```{r}
+mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
+mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')
+
+if (!file.exists(mini_res_dir)) {
   ### MiniMUGA pipeline
@@
   mini_pct_res <- rbindlist(mini_pct_res)
+  saveRDS(minimuga_results, mini_res_dir)
+} else {
+  minimuga_results <- readRDS(mini_res_dir)
+}
+
+if (!file.exists(mini_pct_dir)) {
+  write.csv(mini_pct_res, mini_pct_dir, row.names = F)
+} else {
+  mini_pct_res <- fread(mini_pct_dir)
+}
+
 ```
@@
 ```{r}
-source(paste0(root,"/app.R"))
-shinyApp(ui, server)
+#source(paste0(root,"/app.R"))
+#shinyApp(ui, server)
 ```
@@
 
+##### GigaMUGA samples:
+
 For visualizations, we constructed boxplots that demonstrate the percentage of marker that do not have the same results between qtl2, haploQA and the truth model:
@@
 # filepath to the csvs
-shiny_pct_fp <- file.path(results_dir, 'shiny_pct_csvs')
-
 # DO
@@
 We believe the reason why CC performs so differently is that the optimal model for CC does not allow any heterozygosity, whereas the genail does, so the genail model showed some residual heterozygosity from the CC mice which the optimal model does not have the ability to show.
-
-The outliers at 80% discordance are samples V6 and VP.
-Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate.
-However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous.
-Further investigation may be needed on the two samples.
-
-Below are the karyotype plots for the two samples:
-
-![Karyotype plot - sample VP in CC](/Users/linc/Desktop/VP_geno_cc.png) ![Karyotype plot - sample V6 in CC](/Users/linc/Desktop/V6_geno_cc.png)
 
@@
 
-```{r}
-mean_df <- rbind(t((colMeans(do_pct_df[,2:4]))), t((colMeans(cc_pct_df[,2:4]))), t((colMeans(bxd_pct_df[,2:4]))), t((colMeans(f2_pct_df[,2:4])))) %>% as.data.frame()
-colnames(mean_df) <- paste(colnames(mean_df), "mean", sep = "_")
+For the above three plots, there are two significant outliers at 80% discordance showing for CC, which are samples V6 and VP.
+Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate.
+However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous.
+Further investigation may be needed on the two samples.
 
-mean_df$sample_type <- c('DO', 'CC', 'F3', 'F2')
-mean_df %>% select(sample_type, everything())
-colMeans(cc_pct_df[,2:4])
-```
+Below are the karyotype plots for the two samples:
+
+![Karyotype plot - sample VP in CC](/Users/linc/Desktop/VP_geno_cc.png) ![Karyotype plot - sample V6 in CC](/Users/linc/Desktop/V6_geno_cc.png)
+
+##### Mutant Mice samples:
 
@@
 
+The outlier at approximately 70% is sample JXV, which according to its karyotype on HaploQA, this sample has a significant mount of genetic variations for a mutant mice.
+This could also affect the way qtl2 decides the geno phases and cause the discordance
+
+![Karyotype plot - sample JXV in MiniMUGA](/Users/linc/Desktop/JXV_geno_mini.png)
+
 GigaMUGA mutant:
@@
 ```{r}
-gm_pct_temp <- rbindlist(gm_pct_res)
-gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
-plot_df <- melt(gm_plot)
-plot_df$variable <- 'GigaMUGA Mutant'
+#gm_pct_temp <- rbindlist(gm_pct_res)
+#gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
+#plot_df <- melt(gm_plot)
+#plot_df$variable <- 'GigaMUGA Mutant'
 
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')
+#ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')
 

### 2022-12-16 Scripts/qtl2_sim_comp.R

@@ -2,7 +2,7 @@
 # this is necessary because the computed haploqa results needs to be in the same format as qtl2 (with genoprob, phased geno, etc.)
 ### function directory
 ## 1. maxmarg_sim - simulate maxmarg phasing computations for haploqa
-## 2. save_rds - save all results into rds
+## 2. qtl2_metrics_comp - save all results into rds
 
 
 library(rstudioapi)

### 2022-12-16 haplotype_reconstruction_pipelines.R

@@ -808,7 +808,7 @@ truth_model_reconstruction <- function(sample_type, list_pheno, qtl2_file_gen, s
   }
   
   ### get results for the sample
-  results <- save_rds(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_founders, founder_haplo_lookup, truth_model = T)
+  results <- qtl2_metrics_comp(rds_dir, qtl2_dir, sample_type, results_dir, data_dir, n_founders, founder_haplo_lookup, truth_model = T)
   
   return(results)
 }

### 2022-12-20 draft_report.Rmd
Update draft_report.Rmd

@@
 ---
-title: "Untitled"
+title: "HaploQA vs. R/qtl2 Haplotype Reconstructions"
 output: html_document
-date: '2022-10-18'
+date: '2022-12-16'
 ---
@@
 do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
-
 ```
@@
                           sub = 'Left - qtl2, Right - HaploQA')
+
+# remove DO from memory
 ```
@@
 HaploQA:
+
   1. advantages:
+
       a. Not a programming tool - people without relevant knowledge can still use the tool and perform haplotype reconstruction
+
       b. Works for all MUGA array-based genotyping data
+
       c. Detailed and concise visualizations on web interface
+
   2. disadvantages:
+
       a. haplotype reports are not easy to access computationally
+
       b. source codes are also not easy to access computationally
+
       c. expensive to maintain at the organization's costs
@@
 R/qtl2:
+
   1. advantages:
+
       a. fully accessible computationally
+
       b. capable of performing other genetic analysis
+
       c. currently maintained by developers
+
   2. disadvantages:
+
       a. algorithms is probability based, tends to bounce states when low-contribution founders are present
-      b. does not scale well with large number of founders, run time increases exponentially 
-      c. visualizations are less detailed and need to be generated on demand
+
+      b. visualizations are less detailed and need to be generated on demand
+
+      c. does not scale well with large number of founders, run time increases exponentially
+
+      ![Genoprob runtime plot](/Users/linc/Desktop/genoprob_runtime.png)
 
 #### 5. References
+
 1. http://haploqa-dev.jax.org/help.html

### 2022-12-21 input_data_prep_functions.R
Delete input_data_prep_functions.R

@@ -1,2504 +1,0 @@
-# contains all functions necessary to retrieve data from HaploQA and convert into qtl2 input format
-#
-#
-library(rstudioapi)
-root <- dirname(getSourceEditorContext()$path)
-#
-# function to retrieve summary table
-# @param html_file (html_document) - html of main page sample website on HaploQA, output of read_html(url)
-# @param url_list (list) - url to main page of individuals in said sample, bound on summary table as metadata
-#
-# @return sum_table (data.frame) - table as shown on HaploQA sample website
-# columns of sum_table: ID (sample IDs), secondary ID, Haplotype Candidate (T/F), Strain Name, Sex, % Het Calls, % Hom Calls, % No Call, % Concordance, Sample filepath
-sample_summary_scrape <- function(html_file, url_list, marker_type) {
-  # convert into table
-  html_table <- html_file %>% html_nodes(".table") %>% html_table()
-  sum_table <- html_table[[1]][-1] # remove first column, as it's blank
-  sum_table <- as.data.frame(sum_table, row.names = F) # convert into dataframe
-  # clean up ID columns
-  sum_table$`ID (Secondary IDs)` <- gsub("\\s+", "", sum_table$`ID (Secondary IDs)`) # remove spaces
-  # split IDs and secondary IDs
-  sum_table <- sum_table %>%
-    separate(`ID (Secondary IDs)`, c("ID", "Secondary IDs"), "\\(")
-  sum_table$`Secondary IDs` <- gsub("\\)", "", sum_table$`Secondary IDs`) #remove parentheses
-  sum_table$`Sample Filepath` <- url_list
-  if (!missing(marker_type)) {
-    sum_table <- sum_table[sum_table$Platform == marker_type,]
-  }
-  return(sum_table)
-}
-
-# function to retrieve individuals in a sample
-# @param url (string) - url of main page of each individual 
-# @param url_domain (string) - almost always 'http://haploqa-dev.jax.org/' 
-#
-# @return file (string) - url to direct download of the file associated with individual
-sample_individual_scrape <- function(url, url_domain) {
-  # read html of website with sample data
-  html_sample <- read_html(url)
-  # get class 'btn'
-  sample_list <- html_sample %>% 
-    html_nodes(".btn") %>% html_attr("href")
[…]

### 2022-12-22 draft_report.Rmd
12/22 update

@@
 
+#### 2.1. Environment set-up
+
 The main repository is https://github.com/TheJacksonLaboratory/HaploQA_qtl2_comp which contains the main pipelines developed for this project, all utility functions, and visualizations of selected results.
+
+There is no need to pre-download anything from any other websites or manually create any directories, as those processes were automated by the pipelines.
+However, the automation depend heavily on the CSV files stored in the directory, and without proper set-ups of the CSV files, the pipeline will not run.
+Please make sure the following CSVs are present in the directory after cloning the repo:
+
+1. annotations_config.csv - configuration file containing all technical set-ups for the pipeline, details will be explained in the next section.
+
+2. founder_lookup_table.csv - lookup table containing all possible geno codes (AA, BB, AB, AY, BY, etc.) and their according number codes (1, 2, 3, 4, etc.)
+
+3.
+founder_strains_table.csv - lookup table containing the standard 8 founder strains (A/J, C57BL/6J, etc.) and their according letter codes (A, B, etc.) This is primarily used for CC and DO to ensure the 8 founders are encoded in the correct order.
 
@@
 
-As mentioned in the introduction, in order to decide whether to transition, the accuracy of haploqa and qtl2 each need to be compared to an optimal model.
-The pipelines in the sources script perform haplotype reconstruction tasks using genotyping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, and a GigaMUGA mutant cross), 70 individuals within the MiniMUGA sample.
-Decision was made to not include Muga/MegaMUGA data due to the discontinued usage of the array types.
+There are three major pipelines in the repository:
 
-#### 2.2. Data Retrieval
+a. 'haplotype_reconstruction_pipeline' - haplotype reconstructions for an entire cross (all individuals within the cross)
 
-Input data was retrieved from haploqa.jax.org, and converted into the qtl2 input format. The data wrangling consists of two parts:
-1. scrapes the HTML elements of the haploqa website and downloads the SNP level reports for each individual within each cross
-2. concatenate the individuals within each cross into a single genotype file, converts the genotype file into the required input files for qtl2
+b. 'sample_haplotype_reconstruction' - haplotype reconstructions for one individual
 
-However, haploqa.jax.org contains a large number of uploaded samples and the input data for each cross needs to be retrieved from their according sample.
-For example, for cross 'Collaborative Cross', the URL to the sample is http://haploqa-dev.jax.org/tag/Collaborative%20Cross.html
+c. 'truth_model_reconstruction' - haplotype reconstructions for optimal/truth models
 
-A configuration file, 'annotations_config.csv', is stored in the environment, which allows the pipeline to automatically obtain all relevant information associated with each cross such as the sample URL, input data directory, output data directory, marker annotation file names, etc.
+Some crosses, such as DO and CC, have the same contributing founders across all individuals.
+The haplotype reconstructions can therefore be conducted for all individuals in the entire cross at the same time, and the pipeline 'haplotype_reconstruction_pipeline' can be used for crosses with consistent contributing founders for such purpose.
+Whereas the mutant crosses, such as MiniMUGA, contains individuals with mutate mice where all individuals may have different contributing founder strains.
+Haplotype reconstructions for these types of crosses need to conducted individually for each sample, and therefore, the pipeline 'sample_haplotype_reconstruction' should be used.
+
+To run a cross, make sure to modify the configuration file, 'annotations_config.csv', that's stored in the environment.
+This file allows the pipeline to automatically obtain all relevant information associated with each cross such as the sample URL, input data directory, output data directory, marker annotation file names, etc.
 The purpose of this configuration file is to avoid having to pass a large list of parameters into each pipeline function.
@@
 ```
+
+If the target cross exists in the config file, the pipeline will:
+
+1. get the filename entered in the 'annot_file' column as the annotation file from 'https://raw.githubusercontent.com/kbroman/MUGAarrays/main/UWisc/'
+
+2. get the input data files for all individuals in the cross from the HaploQA url entered in the 'url' column
+
+3. create a folder in the working directory (if not already exists) with the name entered in 'data_dir' column and outputs the scraped input files to this directory
+
+4. create a folder in the working directory (if not already exists) with the name entered in 'qtl2_dir' column to output all the files for qtl2 input
+
+5. get the founder data from the url given in 'founder_url' column
+
+6. run qtl2 with the given value in the 'ngen' column
+
+7. identify the appropriate crosstype to get the optimal model for the given cross using 'truth_model'
+
+8. identify the founders present in the cross. This should be left blank for the mutant crosses or any other crosses where the number of founders ia not consistent between individuals
+
+9. use 'marker_type' to identify the conditions of certain if loops in the pipeline. Should be 'GigaMUGA' or 'MiniMUGA' for the current pipelines
+
+If the cross that the user is running haplotype reconstructions for does not exist in the config file, add a new row and fill each column accordingly.
 
@@
 
-There are three major pipelines in the repository:
-a. 'haplotype_reconstruction_pipeline' - haplotype reconstructions for an entire cross (all individuals within the cross) 
-b. 'sample_haplotype_reconstruction' - haplotype reconstructions for one individual 
-c. 'truth_model_reconstruction' - haplotype reconstructions for optimal/truth models
-
-Some crosses, such as DO and CC, have the same contributing founders across all individuals.
-The haplotype reconstructions can therefore be conducted for all individuals in the entire cross at the same time, and the pipeline 'haplotype_reconstruction_pipeline' can be used for crosses with consistent contributing founders for such purpose.
-Whereas the mutant crosses, such as MiniMUGA, contains individuals with mutate mice where all individuals may have different contributing founder strains.
-Haplotype reconstructions for these types of crosses need to conducted individually for each sample, and therefore, the pipeline 'sample_haplotype_reconstruction' should be used.
+From the above set up, the scripts will look for gm_uwisc_v2.csv in https://raw.githubusercontent.com/kbroman/MUGAarrays/main/UWisc/ as the annotation file.
+The raw input data files will be scraped from http://haploqa-dev.jax.org/tag/Collaborative%20Cross.html and these files will be stored in a folder, 'haploqa_cc_gm', in the same directory as the repo, for further wrangling.
+The wrangled data that's ready for qtl2 read_cross function will be stored in a folder 'cc_qtl2_genail' also in the same directory as the repo.
+qtl2 will be use ngen = 4 to run haplotype reconstructions.
+The optimal model of CC is 'risib8' in qtl2.
 
@@
 
-#### 2.3. Results
+#### 2.2. Results
 
-##### 2.3.1. GigaMUGA
+As mentioned in the introduction, in order to decide whether to transition, the accuracy of haploqa and qtl2 each need to be compared to an optimal model.
+The pipelines in the sources script perform haplotype reconstruction tasks using genotyping data obtained from different types of data sources, specifically, 5 sets of GigaMUGA samples (Collaborative Cross, Diversity Outbred, F3, F2, and a GigaMUGA mutant cross), 70 individuals within the MiniMUGA sample.
+Decision was made to not include Muga/MegaMUGA data due to the discontinued usage of the array types.
+
+##### 2.2.1. GigaMUGA
 
@@
 
-# remove DO from memory
+# remove DO from memory, use gc()
 ```
@@
 ```{r}
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = T)
 ```
@@
 ```{r}
+gm_res_dir <- file.path(results_dir, 'RDS', 'gm_results.rds')
+gm_pct_dir <- file.path(shiny_pct_fp, 'gm_pct_cpmp.csv')
 # ### GigaMUGA mutant
@@
 ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
-ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT')] # only one founder
-
-gigamuga_results <- list()
-gm_pct_res <- list()
+ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT', 'XZ', 'Y9')] # only one founder
+ind_gm <- ind_gm[7:length(ind_gm)]
+#gigamuga_results <- list()
+#gm_pct_res <- list()
+skipped_inds <- list()
 for (ind in ind_gm) {
   print(ind)
-  sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T)
-  df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')
+  skip <- FALSE
+  tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T), error = function(e) {skip <<- TRUE})
+  if(skip) {
+     skipped_inds <- c(skipped_inds, ind)
+     next}
+  #df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')
 
-  gm_pct_res[[ind]] <- df
+  #gm_pct_res[[ind]] <- df
   gigamuga_results[[ind]] <- sample_gm_res
 }
+gigamuga_results <- readRDS(gm_res_dir)
+saveRDS(gigamuga_results, gm_res_dir)
+#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)
 ```
 
-##### 2.3.2. MiniMUGA
+##### 2.2.2. MiniMUGA
 
@@
 
-##### 2.3.3. Shiny Visualizations
+##### 2.2.3. Shiny Visualizations
 
@@
 
-#### 3.Statistical Results
+#### 3. Statistical Results
 
@@
 PMID: 30591514; PMCID: PMC6366910.
+
+Known issues:
+1. Screen time error
+2. Error in dev.off() : 
+  QuartzBitmap_Output - unable to open file '/Users/linc/.local/share/rstudio/notebooks/AFB3D6E3-draft_report/1/8779F85C5B04ECDA/cvdr4ccbjgppi_t/_rs_chunk_plot_010.png'

### 2022-12-22 draft_report.Rmd
updates 12/22

@@
 qtl2_dir <- file.path(root, config_sample$qtl2_dir)
+
+print(config_sample)
 ```
@@
 
-1. Results for DO:
+The pipeline to run haplotype reconstructions for an entire cross can be initiated using the function 'haplotype_reconstruction_pipeline'.
+Please set toggle 'samples_gen' = True if the genotyping data for individuals need to be (re)downloaded from HaploQA, and set to False otherwise.
+please set toggle 'qtl2_file_gen' = F if the input files for qtl2 read_cross need to be (re)generated, and False otherwise.
 
-```{r message=FALSE}
+The function automatically saves the computed haplotype reconstruction results to an RDS file, which you can find in the /results/RDS/ directory of this repo.
+Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance and generate visualizations read the results from these RDS files.
+
+1. Results for DO
+
+```{r, eval=F}
 do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
@@
 
-Below are the genotype visualizations for some individuals:
+Below are the genotype visualizations for some individuals in DO:
 
-```{r}
-plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 1, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+a. Sample 6UY
 
-```{r}
-plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 50, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
-                          sub = 'Left - qtl2, Right - HaploQA')
+![Karyotype plot - sample 6UY in DO](/Users/linc/Desktop/do_plot_1.png)
 
-# remove DO from memory, use gc()
-```
+b. Sample 6WN
+
+![Karyotype plot - sample 6WN in DO](/Users/linc/Desktop/do_plot_2.png)
 
@@
 
-```{r message=FALSE}
+```{r, eval=F}
 cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
@@
 
-Genotype visualizations for some individuals:
+a. Sample 35V
 
-```{r}
-plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 1, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+![Karyotype plot - sample 35V in CC](/Users/linc/Desktop/cc_plot_1.png)
 
-```{r}
-plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 50, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+b. Sample 6MV
+
+![Karyotype plot - sample 6MV in CC](/Users/linc/Desktop/cc_plot_2.png)
 
@@
 
-```{r}
-bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)
+```{r, eval=F}
+bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
 
@@
 
-Genotype visualizations for some individuals:
+a. Sample 5PD
 
-```{r}
-plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 1, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+![Karyotype plot - sample 5PD in F3](/Users/linc/Desktop/bxd_f3_plot1.png)
 
-```{r}
-plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 5, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+b. Sample 5R2
+
+![Karyotype plot - sample 5R2 in F3](/Users/linc/Desktop/bxd_f3_plot2.png)
 
@@
 
-```{r}
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = T)
+```{r, eval=F}
+f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
 ```
 
-Genotype visualizations for some individuals:
+1. Sample 8RS
 
-```{r}
-plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 1, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+![Karyotype plot - sample 8RS in F2](/Users/linc/Desktop/f2_plot1.png)
 
-```{r}
-plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 5, 
-                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
-                          sub = 'Left - qtl2, Right - HaploQA') 
-```
+b. Sample 8T3
+
+![Karyotype plot - sample 8T3 in F3](/Users/linc/Desktop/f2_plot2.png)
 
@@
 ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
-ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
 ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT', 'XZ', 'Y9')] # only one founder
-ind_gm <- ind_gm[7:length(ind_gm)]
+
 #gigamuga_results <- list()
@@
 ```{r}
+
+#haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)
+
 mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
@@
   ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
-  ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
+  ind_mini <- ind_mini[ind_mini!='JXX', 'JY6'] # screentime error, skip this one
   ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
-  ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
+
   minimuga_results <- list()
@@
 cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
-bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
+bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F)
 f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
@@
 
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA - percentage of marker difference between genail and haploqa') + xlab('model')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs qtl2 - percentage of marker discordance') + xlab('model')
 

### 2022-12-22 app.R

@@ -4,7 +4,7 @@ library(data.table)
 
 root <- dirname(getSourceEditorContext()$path)
 
-source("input_data_prep_functions.R", local = T)
+source(paste0(root,"/Scripts/utils.R"), local = T)
 #source(paste0(root,"/util.R"))
 
 ## results directory

### 2022-12-23 draft_report.Rmd
updated documentations

@@
 
-The function automatically saves the computed haplotype reconstruction results to an RDS file, which you can find in the /results/RDS/ directory of this repo.
-Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance and generate visualizations read the results from these RDS files.
+The function automatically saves the computed haplotype reconstruction results to a number of RDS files, which you can find in the /results/RDS/ directory of this repo.
+Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according result RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance metrics and construct visualizations read the results from these RDS files.
 
@@
 bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
-
 ```
@@
 #write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)
+
+for (ind in names(gigamuga_results)) {
+  print(ind)
+  df <- ind_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
+  gm_pct_res[[ind]] <- df
+}
 ```
@@
 
+To summarize the above results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples.
+For MiniMUGA, we selected 25 samples.
+Among which, DO and CC are complex crosses with 8 contributing founders, F3 and F2 are relatively simple two-founder crosses, whereas the number of contributing founders in the mutant crosses samples can vary from 2 to 8.
+This is to ensure there are sufficient evidences to evaluate the accuracy of haplotype reconstruction results for crosses of all computational sizes and complexity.
+
 There are three parts of the comparison:
@@
 As mentioned in the configuration file, the truth model for CC uses crosstype 'risib8', the truth model for DO uses crosstype 'do', the truth model for BXD F3 uses crosstype 'ail', and the truth model for F2 uses crosstype 'f2'.
-The executions of pipelines for the truth models are as shown below:
+The executions of pipelines for the truth models can be initiated using the function, 'truth_model_reconstruction'.
+Same as the previous haplotype reconstruction functions, please set toggle 'samples_gen' = True if the genotyping data for individuals need to be (re)downloaded from HaploQA, and set to False otherwise.
+please set toggle 'qtl2_file_gen' = F if the input files for qtl2 read_cross need to be (re)generated, and False otherwise.
+Please also remove the existing result RDS files for the according cross (if any) whenever the qtl2 input files were regenerated to make sure the results are updated properly.
 
-```{r}
+```{r, eval = F}
 do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
 cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
-bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F)
+bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
 f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)

### 2022-12-23 Scripts/get_comp_results.R

@@ -1,8 +1,16 @@
 ## contains all functions to calculate accuracy metrics and/or to compare results 
 ### function directory
-# 1. location_xo_comp - calculates the distances between crossover locations of haploqa and qtl2 results
-# 2. err_comp
-# 3. comp_df_int
+## 1. location_xo_comp - calculates the distances between crossover locations of haploqa and qtl2 results
+## 2. err_comp
+## 3. comp_df_int
+## 4. genocode_comp_matrix
+## 5. get_geno_pct_diff
+## 6. xo_number_plot
+## 7. loc_xo_distance_plot
+## 8. cos_sim_plot
+## 9. geno_all_comp
+## 10. marker_comp_diff
+
 
 library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
@@ -25,8 +33,7 @@ location_xo_comp <- function(qtl2_pos, haploqa_pos, num_chr) {
     pos_h_chr <- haploqa_pos[[chr]]
     pos_chr <- qtl2_pos[[chr]]
     nmar_h <- sapply(pos_h_chr, length)
-    nmar <- sapply(pos_chr, length) # none has the same length...
-    #x_loc_diff[[chr]] <- list()
+    nmar <- sapply(pos_chr, length) 
     for (i in seq(1:length(pos_chr))) {
       #i <- 149
       x <- unlist(pos_h_chr[i])
@@ -38,7 +45,7 @@ location_xo_comp <- function(qtl2_pos, haploqa_pos, num_chr) {
         df <- apply(abs(outer(y, x, '-')), 2, min)
       }
       x_loc_diff[[chr]][[i]] <- ifelse(is.na(mean(df)), 0, mean(df))
-    } # make a histogram of some sort, 100 bins
+    } 
   }
   return(x_loc_diff)
 }
@@ -48,7 +55,6 @@ err_comp <- function(pr, map, num_chr, results_dir, sample_type) {
   pdf(paste0(results_dir, 'err_comp_plots_', sample_type, '.pdf'))
   facet_all <- data.frame()
   for (chr in num_chr) {
-    #chr <- 2
     print(chr)
     err <- (1 - apply(pr[[chr]], c(1,3), max)) %>% t() %>% as.data.frame()
     err$marker <- rownames(err)
@@ -59,26 +65,21 @@ err_comp <- function(pr, map, num_chr, results_dir, sample_type) {
     
     for (col in seq(2,(ncol(err_chr)-2))) {
       print(col)
[…]

### 2022-12-23 Scripts/qtl2_input_data_prep.R

@@ -1,11 +1,19 @@
 ### Contains all functions to generate qtl2 input data
+### function directory
+## 1. qtl2_geno - reformats raw input data into qtl2 required geno file format
+## 2. qtl2_maps - extracts genetic/physical positions from raw input data into qtl2 required map files format
+## 3. qtl2_pheno - simulates the selected phenotypes (in this case, list_pheno) and reformat them into qtl2 required pheno format
+## 4. qtl2_cov - reformats sample metadata into qtl2 required covariate file format
+## 5. qtl2_foundergeno - reformats raw input data of the contributing founders into qtl2 required geno file format
+## 6. qtl2_ci - reformats more sample metadata into qtl2 required format
+## 7. get_qtl2_input - pipeline function that executes all the above functions and concatenate results 
+## 8. get_founder_data - retrieves raw founder data from haploqa if not exists
 
 library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
 
 source(paste0(root,"/Scripts/input_data_retrieval.R"))
 source(paste0(root,"/Scripts/utils.R"))
-
 
 # function to generate genotype data for qtl2 input
 # @param df (data.frame) - combined dataframe containing all individual files with only necessary columns selected, merged with summary table and chr 0,Y,M and no calls>10% removed
@@ -289,7 +297,6 @@ get_founder_data <- function(founder_url, url_list, sample_type, data_dir, found
     founders_total <- rbindlist(lapply(founder_data_files, read_sample_txt))
     write.csv(founders_total, fp_founders, row.names = F)
   }
-  
 
   return(founders_total)
   
@@ -326,7 +333,6 @@ get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, s
   ## exclude the ones with bad string names
   if(sample_type == 'CC') {
     summary_df <- summary_df %>% filter(grepl('CC', `Strain Name`)) 
-    
   }
   
   # clean up summary df
@@ -409,7 +415,6 @@ get_qtl2_input <- function(data_dir, sample_type, annot_file, qtl2_output_dir, s
     df_founders_encoded <- qtl2_foundergeno(founders_total, founder_url, url_list, founders_dict, annot_encode_df, founders_list, marker_order, founder_haplo_lookup, sample_type)
     file_output[[6]] <- df_founders_encoded 
   }
-  
   
   print('founder geno data done')
   

### 2022-12-23 Scripts/qtl2_sim_comp.R

@@ -1,13 +1,12 @@
 # contains all functions that aligns haploqa results with that of qtl2 and save the results
 # this is necessary because the computed haploqa results needs to be in the same format as qtl2 (with genoprob, phased geno, etc.)
+
 ### function directory
 ## 1. maxmarg_sim - simulate maxmarg phasing computations for haploqa
-## 2. qtl2_metrics_comp - save all results into rds
-
+## 2. qtl2_metrics_comp - computes qtl2 metrics (i.e. genoprob and phased geno) and save all results into rds
 
 library(rstudioapi)
 root <- dirname(getSourceEditorContext()$path)
-
 
 source(paste0(root,"/Scripts/utils.R"))
 
@@ -44,33 +43,15 @@ maxmarg_sim <- function(summary_df, data_dir, num_chr, founder_all_rev_lookup, c
     ifelse(rev_col %in% founder_all_rev_lookup, rev_col, col)
   })
   
-  #pr <- cross[['pr']]
-  
-  ### join sex into this df - for males, take first letter, then paste Y after it
-  #founder_all_lookup <- setNames(names(founder_all_rev_lookup), founder_all_rev_lookup) # reverse the lookup table
-  #founder_all_lookup <- founder_all_lookup[names(founder_all_lookup) %in% unique(haploqa_maxmarg$haplotype)]
-  #founder_all_lookup <- setNames(seq(1, length(founder_all_lookup)), names(founder_all_lookup))
-  #haploqa_maxmarg[,4] <- as.data.frame(apply(haploqa_maxmarg[,c(4)], 2, function(x) founder_all_lookup[x]))
-  ### subset the lookup table names to just what is in haploqa_maxmarg$haplotype
-  ### then assign numbers seq(1, length())
-  
-  
   ## save per chromosome
   ginf_haploqa <- list()
   for (i in num_chr) {
-     #i <- '1'
-    
     # founder table
     geno_codes <- colnames(pr[[i]])
-    #founder_all_lookup <- setNames(names(founder_all_rev_lookup), founder_all_rev_lookup)
     founder_all_lookup <- setNames(seq(1, length(geno_codes)), geno_codes)
-    print(founder_all_lookup)
-    
     haploqa_maxmarg_chr <- haploqa_maxmarg %>% filter(chromosome == i)
-    
[…]

### 2022-12-23 Scripts/utils.R

@@ -1,5 +1,12 @@
 ### Contains all general utility functions used across the board
-
+### function directory
+## 1. get_haplotypes
+## 2. sort_chr - sort chromosomes from 1-19 and X as needed
+## 3. make_cross - adds crosstype attributes to an object to make it a cross (as needed)
+## 4. shiny_viz_input
+## 5. get_raw_geno - 
+## 6. plot_onegeno_test - modified version of plot_onegeno in qtl2, plots two genos instead of one
+## 7. addgenorect - util function that adds rectangles based on geno positions, used in plot_onegeno_test
 
 
 # function to retrieve only the dataframe that all qtl2 input dataframes were based on for testing purposes, and/or the haplotype columns
@@ -20,7 +27,6 @@ get_haplotypes <- function(summary_df, data_dir) {
   sum_df$`% No Call` <- as.numeric(gsub("%", "", sum_df$`% No Call`))
   
   # eliminate those that has no call > 10%
-  ### use checkifnot to make sure the markers/SNP are in the annotation file (annotation file is the boss)
   df_all <- merge(df_raw, sum_df, by = 'sample_id') %>% filter(`% No Call` < 10) %>% filter(!chromosome %in% c('0', 'Y', 'M'))
   
   return(df_all)
@@ -229,7 +235,6 @@ plot_onegeno_test <- function(geno, geno1, map, ind=1, chr=NULL,
         axis(side=2, at=pretty(ylim), mgp=mgp.y, las=las, tick=FALSE)
       }
       
-      
       # grid lines
       if(!(length(vlines)==1 && is.na(vlines))) {
         abline(v=vlines, col=vlines_col, lwd=vlines_lwd, lty=vlines_lty)
@@ -327,9 +332,6 @@ plot_onegeno_test <- function(geno, geno1, map, ind=1, chr=NULL,
              x_lower_right, max(map[[i]], na.rm=TRUE),
              col=NULL, border=border, lend=1, ljoin=1)
         inc <- inc + 1.5
-        
-        
-        
       }
       
       ## plot x individually
@@ -343,9 +345,7 @@ plot_onegeno_test <- function(geno, geno1, map, ind=1, chr=NULL,
       
       inc <- inc + 1
       
-      
       if(is.matrix(g)) {
-        print('x has only one')
         ### dataframe - g
         x_lower_left <- i+0.75+inc
[…]

### 2022-12-23 haploqa_qtl2_report.Rmd
Updating report

--- 2022-12-23 draft_report.Rmd
@@
 ```{r}
-
 library(rstudioapi)
@@
 
+The below two code chunks represent how the pipeline is used for the mutant mice crosses, where the contributing founders vary between crosses.
+The process involve a certain extent of filtering because some samples within the cross may only have one contributing founder, which does not work for genAIL as the algorithm requires two or more founders to perform haplotype reconstructions, and there is no simple way of identifying how many contributing fouders are present in each individual.
+
+The function below again, saves the RDS files that contain haplotype reconstruction results for all samples in the same RDS directory as above (/results/RDS/), please make sure to remove th existing RDS file if the results need to be regenerated.
+The haplotype reconstruction results for each individuals can also be found in the directory where the qtl2 input files for read_cross are being stored - which should be the specified 'qtl2_dir' value in the config file.
+
 5. GigaMUGA Mutant cross
 
-```{r}
+```{r, eval=F}
 gm_res_dir <- file.path(results_dir, 'RDS', 'gm_results.rds')
-gm_pct_dir <- file.path(shiny_pct_fp, 'gm_pct_cpmp.csv')
+gm_pct_dir <- file.path(shiny_pct_fp, 'gm_pct_comp.csv')
+
+if (!file.exists(gm_res_dir)) {
   # ### GigaMUGA mutant
@@
 
-#gigamuga_results <- list()
-#gm_pct_res <- list()
+  gigamuga_results <- list()
+  gm_pct_res <- list()
   skipped_inds <- list()
@@
     skip <- FALSE
-  tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T), error = function(e) {skip <<- TRUE})
+    tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = F, qtl2_file_gen = F), error = function(e) {skip <<- TRUE})
     if(skip) {
@@
        next}
-  #df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')
+    df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')
 
-  #gm_pct_res[[ind]] <- df
+    gm_pct_res[[ind]] <- df
     gigamuga_results[[ind]] <- sample_gm_res
   }
+  print(paste0('Skipped individuals:', skipped_inds))
+  gm_pct_res <- rbindlist(gm_pct_res)
+  saveRDS(gigamuga_results, gm_res_dir)
+} else {
   gigamuga_results <- readRDS(gm_res_dir)
-saveRDS(gigamuga_results, gm_res_dir)
+}
+
+if (!file.exists(gm_pct_dir)) {
+  write.csv(gm_pct_res, gm_pct_dir, row.names = F)
+} else {
+  gm_pct_res <- fread(gm_pct_dir)
+}
+
 #write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)
 
-for (ind in names(gigamuga_results)) {
+for (ind in names(gigamuga_results)[21:100]) {
   print(ind)
@@
 
-```{r}
-
-#haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)
-
+```{r, eval=F}
 mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
@@
   mini_pct_res <- list()
+  skipped_inds <- list()
   for (ind in ind_mini) {
-    #ind <- 'JXN'
     print(ind)
-    sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
+    skip <- FALSE
+    tryCatch(sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F), error = function(e) {skip <<- TRUE})
+    if(skip) {
+       skipped_inds <- c(skipped_inds, ind)
+       next}
     df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')
@@
   }
-
+  print(paste0('Skipped individuals:', skipped_inds))
   mini_pct_res <- rbindlist(mini_pct_res)
@@
 
-```{r}
-#source(paste0(root,"/app.R"))
-#shinyApp(ui, server)
+A dashboard was built in R Shiny to visualize the karyotype plot results of the haplotype reconstructions, the script to the Shiny app is 'app.R' in this directory.
+The below code chunk should be able to launch the Shiny app.
+
+```{r, eval = F}
+source(paste0(root,"/app.R"))
+shinyApp(ui, server)
 ```
@@
 
-To summarize the above results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples.
+To summarize the results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples.
 For MiniMUGA, we selected 25 samples.
@@
 f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
+# GigaMUGA Mutant
+gm_pct_res <- fread(file.path(shiny_pct_fp, 'gm_pct_comp.csv'))
+# MiniMUGA
+mini_pct_res <- fread(file.path(shiny_pct_fp, 'mini_pct_comp.csv'))
 ```
@@
 
-ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs qtl2 - percentage of marker discordance') + xlab('model')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs genail - % of marker discordance') + xlab('model')
 
@@
 ```{r}
-#gm_pct_temp <- rbindlist(gm_pct_res)
-#gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
-#plot_df <- melt(gm_plot)
-#plot_df$variable <- 'GigaMUGA Mutant'
+gm_plot <- gm_pct_res %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
+plot_df <- melt(gm_plot)
+plot_df$variable <- 'GigaMUGA Mutant'
 
-#ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')
+ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant haploqa vs genail - % of marker discordance') + xlab('model')
 

### 2022-12-23 haploqa_qtl2_report.Rmd
12/23 updates

@@
 }
+```
 
-#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)
+The selected GigaMUGA mutant cross contains some mice with two contributing founders and some genetically diverse DO mice:
 
-for (ind in names(gigamuga_results)[21:100]) {
-  print(ind)
-  df <- ind_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
-  gm_pct_res[[ind]] <- df
-}
-```
+1. Sample ND
+
+![Karyotype plot - sample ND in GigaMUGA Mutant](/Users/linc/Desktop/gm_plot_ND.png)
+
+b. Sample YE
+
+![Karyotype plot - sample YE in GigaMUGA Mutant](/Users/linc/Desktop/gm_plot_YE.png)
 
@@
 ```
+
+a. Sample JXN
+
+![Karyotype plot - sample JXN in MiniMUGA](/Users/linc/Desktop/mini_plot_jxn.png)
+
+b. Sample JYA
+
+![Karyotype plot - sample JYA in MiniMUGA](/Users/linc/Desktop/mini_plot_jya.png)
 
@@
 ```{r}
+# wrangle csv files for plots
 do_qtl2_truth_diff <- do_pct_df$qtl2_pct_diff
@@
 
+The results between genail and haploQA appear to be highly similar, with an average of less than 2% marker discordance between the models.
+
 #### 4. Conclusion
@@
 Known issues:
-1. Screen time error
+
+1.
+The webpages throws a screentime error while downloding some individuals from haploQA - this is presumably a Mac issue.
+The functions automatically skip the samples with screentime error issue, however, this may become an issue when a screentime error rises up while getting the founder data because the function would skip that founder, and the user needs to manually download it.
+Please make sure that all founders are present before proceeding with the pipelines.
+
 2. Error in dev.off() : 
   QuartzBitmap_Output - unable to open file '/Users/linc/.local/share/rstudio/notebooks/AFB3D6E3-draft_report/1/8779F85C5B04ECDA/cvdr4ccbjgppi_t/_rs_chunk_plot_010.png'
+
+This is likely an R markdown issue or some unclosed connections in the backend environment after knitting the R markdown - please make sure to clear the workspace or restart R if this happens.

### 2022-12-23 HaploQA_qtl2_output.R

@@ -1,371 +1,0 @@
-##### HaploQA data reformatting - Launch script
-### used for qtl2 input
-### annotation config saved in annotations_config.csv
-### founder strain dictionary saved in founder_strains_table.csv
-
-### output files for haplotype reconstruction
-
-
-## everything that starts from BJSJL
-
-# take one founder out (cast), use genail 7
-# 6UY DO
-# run reconstruction, plot onegeno and run error lod
-# find in python where haploqa generates the plots
-
-# 1. DO genail 7, do error lod computations
-# 2. document
-# 3. haploqa codes
-# 4. 
-sample_res <- sample_haplotype_reconstruction('MURGIGV01', 'PX', samples_gen = F, qtl2_file_gen = T)
-
-
-library(rstudioapi)
-library(qtl2)
-library(dplyr)
-library(tidyverse)
-library(httr)
-library(rvest)
-library(data.table)
-library(qtl2convert)
-library(ggrepel)
-library(reshape2)
-library(stats)
-library(ggplot2)
-library(stringr)
-library(stringi)
-library(lsa)
-library(readr)
-
-# 2, 4, 8, 16 founders
-# pipeline for one GigaMUGA mutant sample
-# system.time()
-
-# source the function script
-root <- dirname(getSourceEditorContext()$path)
-source(paste0(root,"/input_data_prep_functions.R"))
-
-# results directory
-results_dir <- file.path(root, 'results')
-dir.create(results_dir, showWarnings = FALSE) 
-
-# simulated geno
-list_pheno <- c('WBC', 'NEUT')
-
-### GigaMUGA pipelines
-do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
-cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
-bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
-f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
-
-## GigaMUGA truth models
[…]

### 2022-12-23 Scripts/get_comp_results.R

@@ -1,15 +1,15 @@
 ## contains all functions to calculate accuracy metrics and/or to compare results 
 ### function directory
 ## 1. location_xo_comp - calculates the distances between crossover locations of haploqa and qtl2 results
-## 2. err_comp
+## 2. err_comp 
 ## 3. comp_df_int
 ## 4. genocode_comp_matrix
 ## 5. get_geno_pct_diff
 ## 6. xo_number_plot
 ## 7. loc_xo_distance_plot
-## 8. cos_sim_plot
-## 9. geno_all_comp
-## 10. marker_comp_diff
+## 8. cos_sim_plot - plots the cosine similarities
+## 9. geno_all_comp - reformats the alleles of haploqa, genail and truth model
+## 10. marker_comp_diff - computes the percentage of marker discordance based on the reformatted alleles of geno_all_comp
 
 
 library(rstudioapi)

### 2022-12-23 Scripts/utils.R

@@ -1,10 +1,10 @@
 ### Contains all general utility functions used across the board
 ### function directory
-## 1. get_haplotypes
+## 1. get_haplotypes - retrieve only the dataframe that all qtl2 input dataframes were based on for testing purposes
 ## 2. sort_chr - sort chromosomes from 1-19 and X as needed
 ## 3. make_cross - adds crosstype attributes to an object to make it a cross (as needed)
-## 4. shiny_viz_input
-## 5. get_raw_geno - 
+## 4. shiny_viz_input - gets only the input needed for the Shiny app, primarily used for easy access in app.R
+## 5. get_raw_geno - gets only the raw geno alleles, used for calculations of marker discordance
 ## 6. plot_onegeno_test - modified version of plot_onegeno in qtl2, plots two genos instead of one
 ## 7. addgenorect - util function that adds rectangles based on geno positions, used in plot_onegeno_test
 
